### 2016-06-15 doc/NEAMS_Report/NEAMS_Report.tex
final review.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[12pt]{scrartcl} 
%########################### Preferences #################################

\usepackage{vmargin}
\usepackage{color}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amstext}
\usepackage{amsbsy}
%\usepackage{mathbbol}
\usepackage{graphicx} 
\usepackage{verbatim}
\usepackage{csvsimple} 
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{fancyhdr}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{listings}
\usepackage[numbers,sort&compress]{natbib}[2010/09/13]

%  new definitions
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Equation~\ref{#1}}                     % equation
\newcommand{\fig}[1]{Figure~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\bs}[1]{\mathbf{#1}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\tcb}[1]{\textcolor{blue}{#1}}

% tikz stuff
\usetikzlibrary{shapes,arrows,positioning}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]
\newcommand{\tikzback}[5]{
 \begin{pgfonlayer}{background}
  \path (#1.west |- #2.north)+(-0.5,0.75) node (a1) {};
  \path (#3.east |- #4.south)+(+0.5,-0.25) node (a2) {};
  \path[fill=yellow!10,rounded corners, draw=black!100, dashed] (a1) rectangle (a2);
   \path (#3.east |- #2.north)+(0,1.25)--(#1.west |- #2.north) node[midway] (#5-n) {};
   \path (#3.east |- #2.south)+(0,-0.35)--(#1.west |- #2.south) node[midway] (#5-s) {};
   \path (#1.west |- #2.north)+(-0.75,0)--(#1.west |- #4.south) node[midway] (#5-w) {};
 \end{pgfonlayer}}

% input stuff
\lstset{%
%  language=C++,
  showstringspaces=false,
% basicstyle=\scriptsize\ttfamily,
  basicstyle=\small\ttfamily,
  commentstyle=\color{inl@green},
%  keywordstyle=\bfseries,
%  escapeinside=$$
%  framesep=0pt,
%  rulesep=0pt,
  frame=single
}
% Make nice big tildes inside lstlistings instead of
% tiny ones.
\lstset{
    literate={~} {$\sim$}{1}
}

% ********* Caption Layout ************
%\usepackage{ccaption} % allows special formating of the captions
%\captionnamefont{\bf\footnotesize\sffamily} % defines the font of the caption name (e.g. Figure: or Table:)
%\captiontitlefont{\footnotesize\sffamily} % defines the font of the caption text (same as above, but not bold)
%\setlength{\abovecaptionskip}{0mm} %lowers the distance of captions to the figure
\setlength\parindent{0pt}
\setlength{\oddsidemargin}{1in}
\setlength{\evensidemargin}{1in}
\setlength{\textwidth}{6.5in}
% ********* Header and Footer **********
% This is something to play with forever. I use here the advanced settings of the KOMA script

\title{Improved Quasi-Static Methods in Rattlesnake}
\subtitle{Development and Implementation of Improved Quasi-Static (IQS) methods for time-dependent neutron diffusion and neutron transport solvers in Rattlesnake}
\author{ \normalsize
  \textbf{Zachary M. Prince$^\dagger$, Jean C. Ragusa$^\dagger$, Yaqi Wang$^\star$} \\
 \normalsize $^\dagger$Department of Nuclear Engineering \\
 \normalsize Texas A\&M University, College Station, TX, USA\\
 \normalsize $^\star$Idaho National Laboratory\\
 \normalsize \href{mailto:zachmprince@tamu.edu}{zachmprince@tamu.edu}, \href{jean.ragusa@tamu.edu}{jean.ragusa@tamu.edu}, \href{yaqi.wang@inl.gov}{yaqi.wang@inl.gov} 
}

\hypersetup{
  colorlinks=true,
  urlcolor=blue,
}

\pagestyle{fancy}
\fancyhf{}
\lhead{INL/EXT-16-38059}
\rfoot{DRAFT}
\cfoot{\thepage}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% boxes
\usepackage{color}
\definecolor{myblue}{rgb}{.8, .8, 1}
\usepackage{empheq}

\newlength\mytemplen
\newsavebox\mytempbox

\makeatletter
\newcommand\mybluebox{%
    \@ifnextchar[%]
       {\@mybluebox}%
       {\@mybluebox[0pt]}}

\def\@mybluebox[#1]{%
    \@ifnextchar[%]
       {\@@mybluebox[#1]}%
       {\@@mybluebox[#1][0pt]}}

\def\@@mybluebox[#1][#2]#3{
    \sbox\mytempbox{#3}%
    \mytemplen\ht\mytempbox
    \advance\mytemplen #1\relax
    \ht\mytempbox\mytemplen
    \mytemplen\dp\mytempbox
    \advance\mytemplen #2\relax
    \dp\mytempbox\mytemplen
    \colorbox{myblue}{\hspace{1em}\usebox{\mytempbox}\hspace{1em}}}

\makeatother

%################ End Preferences, Begin Document #####################

\pagestyle{plain} % on headers or footers on the first page

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% set relative directory for figures. can be changed at any time later with \renewcommand
\newcommand{\FiguresDir}{./figs}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\maketitle
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\paragraph*{Abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This report presents  the implementation of the Improved Quasi-Static method (IQS) in the Yak/Rattlesnake applications of the MOOSE framework.  In the IQS method, the neutron flux is factored into a time-dependent amplitude and a spatial- and weakly time-dependent shape.  The shape evaluation is very similar to a standard flux solve and is computed at large (macro) time steps. The amplitude is obtained by solving the PRKEs on a micro time scale. The PRKE parameters are regularly updated based on the shape changes.  Two versions of the IQS methodology have been implemented: (i) the standard approach, which leads to a system of nonlinear equations between the shape and the amplitude, and (ii) a predictor corrector approach that linearizes the system of equations. A time-adaptation technique has been developed to automatically select an appropriate macro time step size based on user-specified tolerances. The implementation has been verified and tested using manufactured solutions, a custom one-dimensional example, the TWIGL ramp benchmark, and the LRA benchmark.  The numerical results include error convergence plots and performance with time adaptation.  These examples prove it to be a viable and effective method for  transient neutronic simulations. More complex simulations (e.g., TREAT transients) are planned to perform coupled neutronics/heat conduction transient simulations within the MAMMOTH application.

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\par
The improved quasi-static (IQS) method is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (the shape) and a time-dependent component (the amplitude) \cite{Ott_1966,Dulla2008}. The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates. The IQS method has primarily been developed in the context of the neutron diffusion approximation \cite{Ban_2012,Ferguson_1973}, but reference \cite{Goluoglu_2001} provides an extension of the technique to transport, implemented in TDTORT. The method has mostly being applied to neutron kinetics, without thermal-hydraulic feedback. In this report, we extend the IQS method in the context of transient multiphysics simulations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS for Multigroup Diffusion Equations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this Section, we recall the equations for the IQS method, using the standard multigroup diffusion equations written below:
\begin{subequations}
\begin{align}
\frac{\partial}{\partial t}\left(\frac{\phi^g}{v^g}\right)  =& \frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \phi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g \right) \phi^g  \nonumber \\
&  + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \phi^{g'}  + \sum_{i=1}^I\chi_{d,i}^g\lambda_i C_i \ , \quad 1 \le g \le G 
\label{eq:flux}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\sum_{g=1}^G\nu^{g} \Sigma_f^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee
\end{subequations}
%
Extension to the transport equations will be provided subsequently, once a concise operator notation has been introduced. 
In the above equations,  \\

\begin{tabular}{lll}
$\phi^g$   			&	$=$	&	Scalar flux in energy group $g$ \\
$C_i$					  &	$=$	&	Concentration of delayed neutron precursor $i$ \\
$\Sigma_f^{g}$	&	$=$	&	Fission cross section in energy group $g$ \\
$\Sigma_r^{g}$	&	$=$	&	Removal cross section in energy group $g$ \\
$\Sigma_s^{g' \to g}$	&	$=$	&	Scattering cross section from energy group $g'$ to $g$ \\
$v^g$					  &	$=$	&	Neutron velocity in energy group $g$ \\
$\chi_p^g$			&	$=$	&	Fission spectrum of prompt neutrons \\
$\chi_{d,i}^g$	&	$=$	&	Fission spectrum of delayed neutrons from precursor $i$ \\
$\nu^g$					&	$=$	&	Total number of neutrons per fission \\
$D^g$					  &	$=$	&	Diffusion coefficient in energy group $g$\\
$\lambda_i$			&	$=$	&	Decay constant of precursor $i$ \\
$\beta_i$				&	$=$	&	Delayed neutron fraction  from precursor $i$ \\
$\beta$			 	  &	$=$	&	Total delayed neutron fraction ($\beta = \sum_{i=1}^I \beta_{i}$) \\
  & & 
\end{tabular}

Note that the prompt fission terms have been normalized by $\keff$, the eigenvalue of the initial state, as transient simulations are often carried out for initially critical reactors. The expressions for subcritical and source-driven problems are straightforward to derive but omitted in this report for brevity. \\

The factorization approach in IQS leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent multigroup shape ($\varphi^g$):
\be
\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
\ee
We stress that the factorization is not an approximation. Reporting it in the flux and precursor equations yields, after some simple algebra, the following equations for the shape and the precursors:
\begin{subequations}
\begin{align}
\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right) = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
& -  \left( -\div D^g \grad  + \Sigma_r^g + \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right) \varphi^g + \boxed{\frac{1}{p}}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
\label{eq:shape2}
\end{align}
\be
\frac{dC_i}{dt} = \frac{\beta_i}{\keff}\boxed{p} \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:preq2}
\ee
\end{subequations}

We note that the time-dependent shape equation is similar to the time-dependent flux, with the following modifications:
\begin{enumerate}
\item The shape equation contains an additional term equivalent to a removal cross section,  $ \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$.
\item The delayed neutron source term is divided by $p$.
\item The system of equations is now nonlinear due to the factorization.
\item An equation is needed to obtain the amplitude $p$.
\end{enumerate}

To derive the amplitude equation (also known as the PRKE equations), the shape/precursors equations are weighed by a space-dependent function (typically the initial adjoint flux, $\phi^{*g}$) and integrated over the phase-space. The final expressions are given below:
\begin{subequations}
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\label{eq:p}
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\label{eq:c}
\ee
\end{subequations}
where we have defined the reactivity, effective delayed-neutron fraction, and delayed-neutron precursor decay constant as follows:
\begin{subequations}
\be
\frac{\rho-\bar{\beta}}{\Lambda} = 
\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}
{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda} = \sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda} = 
\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}
{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:b}
\ee
\be
\bar{\lambda}_i = \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee 
\end{subequations}
The following inner product definition has been used: $\left(\phi^{*g},f\right):=\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r$.\\  
Additionally, in order to impose uniqueness on the factorization and to derive the PRKE, the following normalization condition is imposed: $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right) = \textit{constant}$.  This condition is later used  as the convergence criteria when solving the nonlinear system of equations resulting from the IQS method.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Operator Notation and Extension of IQS to Multigroup Transport Equations}
\label{sect:transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For simplicity, the previous section only described the IQS implementation using the neutron diffusion equation. However, deriving the IQS form for other neutron balance equations (e.g., transport, simplified transport, etc.) is very similar. To this end, we re-write the neutron conservation equations in operator form, shown in \eqt{eq:transport}. 
\be
\frac{\partial}{\partial t}\left(\frac{\Psi^g}{v^g}\right) = \sum_{g'} \left(H^{g'\to g} + P_p^{g' \to g} \right) \Psi^{g'} - L^g\Psi^g + S_d^g
\label{eq:transport}
\ee
where $\Psi^g$ is the multigroup neutron flux (angular flux in the case of transport), $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt neutron production operator, $L^g$ is the loss operator, and $S_d^g$ is the delayed neutron source operator. Using \eqt{eq:flux}, the reader may easily obtain the functional form for these operators in the case of a diffusion approximation.
%
%Factorization is defined as $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ and the resulting shape equation is defined by \eqt{eq:tshape}.  
Next, the factorization $\Psi^g(\vec{r},\vec{\Omega},t)=p(t)\psi^g(\vec{r},\vec{\Omega},t)$ is introduced, where the shape is denoted by $\psi^g$, leading to the following shape equations, \eqt{eq:tshape}:
\be
\frac{\partial}{\partial t}\left(\frac{\psi^g}{v^g}\right) = \sum_{g'} \left(H^{g' \to g} + P_p^{g' \to g}\right) \psi^{g'} - \left( L^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \psi^g + \frac{1}{p} S_d^g
\label{eq:tshape}
\ee
Finally, the PRKE parameters are defined by Equations~\ref{eq:trmb}~and~\ref{eq:tb}, where 
$\left(\Psi^{*g},f^g\right) =$\\ $ \int_{4\pi}\int_D \Psi^{*g}(\vec{r},\vec{\Omega})f^g(\vec{r},\vec{\Omega})d^3r d^2\Omega$.
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\Psi^{*g},\sum_{g'}(H^{g' \to g}g+P_p^{g' \to g}-L^{g'}\delta_{g'g})\psi^{g'}\right)}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
\label{eq:trmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\Psi^{*g}, \sum_{g'} P_{d,i}^{g' \to g} \psi^{g'})}{\sum_{g=1}^G\left(\Psi^{*g},\frac{1}{v^g}\psi^g\right)}
\label{eq:tb}
\ee
where $P_{d,i}^{g' \to g}$ is the delayed-neutron operator for precursor group $i$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Solving the IQS System of Equations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Solving for the shape can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape. The PRKE equations form a small ODE system and can be solved using a much smaller time step size. In transients where the shape varies much less than the flux, IQS can thus be very computationally effective. The two time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.

\begin{figure}[!htbp]
\includegraphics[width=\linewidth]{\FiguresDir/IQS_visualization.jpg}
\caption{IQS method visualization}
\label{fig:iqsviz}
\end{figure}

By this design, the benefits of IQS is most realized when using time adaptive techniques, one of these is described in  \sct{sect:dt2}.  As we noted earlier, shape-PRKE equations are a nonlinear system and thus need to be solved in a iterative manner.  Each macro time step can be iterated so the best shape is used to compute power at the micro time steps.  This iteration process must converge the shape such that the uniqueness condition $(\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=0)$ is preserved.

The most convenient method to implement IQS in RATTLESNAKE is through Picard or fixed-point iterations. Over a given Picard iteration, the shape is evaluated first with the latest available amplitude. The updated shape is employed to compute the PRKE parameters (using a linear interpolation in time) and then the PRKEs are solved using micro time steps. This process is repeated until convergence of the normalization condition. Picard iterations can also be utilized for multiphysics simulations, where neutronics (e.g., IQS) is nonlinearly coupled to thermal-hydraulics, for instance.  Thus, inlcuding the IQS nonlinear system solve into the Picard iteration of MOOSE is a straightforward process and is described in \fig{fig:picard}.

\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Solve shape equations \\ over macro time step};
\node[greenblock](p2) at (0,-2.5) {Compute PRKE parameters \\ over macro time step};
\node[greenblock](p3) at (0,-5) {Solve PRKE using micro \\ steps};
\node[greenblock](p4) at (0,-7.5) {Update $p$ and $\frac{dp}{dt}$};
\node[blueblock] (p5) at (0,-11) {Solve other physics \\ components};
\node[reddiamond] (check) at (0,-15) {Check for \\ convergence};
\node [above =0.1mm of p1] {\large{IQS Solve:}};
\node [above =0.1mm of p5] {\large{Multi-physics:}};

\tikzback{p1}{p1}{p4}{p4}{bk1}
\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- (p2.north);
\draw[->,ultra thick](p2.south) -- (p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)+(0,-0.25) -- node [above] {} (bk2-n);
\draw[->,ultra thick](p5.south)+(0,-0.25) -- (check.north);
%\draw[->,ultra thick](check.west) -- (-5,-15) -- node[above,sloped] {\large{no}}(-5,0) -- node[above] {}(bk1-w);
\draw[->,ultra thick](check.west) -- node[above,sloped] {\large{no}} (-5,-15) |-  node[above] {}(bk1-w);
\draw[->,ultra thick](check.south) -- node[right] {\large{yes}} ++(0,-2);

\end{tikzpicture}
\caption{Visualization of Picard iteration process}
\label{fig:picard}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Predictor-Corrector version of IQS (IQS P-C)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The Predictor-Corrector (P-C) version of IQS factorizes the flux and derives the PRKE the same way as the standard version, but the solution of the coupled system of equations is different.  In the IQS P-C version, the flux equations (not the shape equations) are solved (represented by Equations~\ref{eq:flux}~and~\ref{eq:precursor}) in order to obtain a predicted flux solution. This predicted flux is then converted to shape by rescaling as follows:
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
where the scaling factors are given by
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0} \,.\right)
\ee

The PRKE parameters are then computed with this shape using Equations~\ref{eq:rmb}~-~\ref{eq:l} and interpolated over the macro step, then the PRKE ODE system is solved on the micro time scale.  With the newly computed amplitude, the shape is rescaled into a flux and the final corrected flux is given by:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
\ee

The advantage to the predictor-corrector method is there is no iteration necessary for this method and, in turn, is much simpler and faster than the standard IQS.  
%The disadvantage is that the method assumes, by construction, that $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)$ is inherently constant.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Precursor Integration}
\label{sect:prec}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section presents the time-integration method used to solve coupled IQS shape and precursor equations, represented by Equations~\ref{eq:shape2}~and~\ref{eq:preq2}. First, we note that we could keep this system of two time-dependent equations and solve it as a coupled system. However, this is unnecessary and a memory expensive endeavor because the precursor equation is only an ODE and not a PDE. Instead, one may discretize in time the shape equation, which typically requires the knowledge of the precursor concentrations at the end of the time step. This precursor value is taken from the solution, numerical or analytical, of the precursors ODE.  \\

Through prototyping, it has been found that typical implicit nor Crank-Nicholson time discretizations of precursors are not preferable methods for solving the shape equation in IQS.  It has been found that these discretizations result in a lack of convergence of the shape over the IQS iteration process.  In order to remedy the error, a analytical representation of the precursors was implemented in the prototype and the shape solution was able to converge.  The method entails linearly interpolating the fission source ($S_f = \nu \Sigma_f \varphi$) and integrating the resulting continuous function analytically. \eqt{eq:c_dis} is the equation to solve for the precursors and Equations~\ref{eq:a1}~and~\ref{eq:a2} are integrating factors.

\be
C^{n+1} = C^n e^{-\lambda \Delta t} 
+ \beta_i \left(a_2 S_f^{n+1} + a_1 S_f^{n}\right)
\label{eq:c_dis}
\ee

\be
a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
\label{eq:a1}
\ee
\be
a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')e^{-\lambda(t_{n+1}-t')}dt'
\label{eq:a2}
\ee

The amplitude $(p)$ is included in the integration coefficient because it has been highly accurately calculated in the micro step scheme, so a piecewise interpolation between those points can be done to maximize accuracy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Time Adaptation}
\label{sect:dt2}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

IQS aims at reducing the time discretization error in the flux solution by splitting the flux into an amplitude (that is highly resolved with at a micro time scale) and a shape (whose time-dependence is weaker than that of the flux itself). Thus, by construction, the IQS approach may employ larger time-step sizes for comparable temporal error. Further enhancements can be gained by using time adaptation (or time step control) in order to increase or reduce the macro time step size for the
shape evaluation, depending on error estimates. Here, we employ a step-doubling technique to carry out the time adaptation.
The step doubling technique involves estimating the local error for a certain time step by taking the difference between a solution with one full step ($\varphi^g_{\Delta t}$) and a solution with two half steps ($\varphi^g_{\Delta t/2}$). Note:
$\varphi$ is changed to $\phi$ for regular flux evaluation (brute-force time discretization of the governing equations) and IQS P-C.

The relative error is computed as follows:
\be
e_n = \frac{\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2} - \sum_{g=1}^G\varphi^g_{\Delta t}}}{\text{max}\left(\norm{\sum_{g=1}^G\varphi^g_{\Delta t/2}},\norm{\sum_{g=1}^G\varphi^g_{\Delta t}}\right)}
\label{eq:edt2}
\ee
If the error is smaller than the user-specified tolerance, $e_{tol}$, the time step is accepted. In addition, a new time step size is estimated as follows:
\be
\Delta t_{new} = S \Delta t \left(\frac{e_{tol}}{e_n}\right)^{\frac{1}{1+q}} \,,
\label{eq:dt2}
\ee
where $q$ is the convergence order of the time integration scheme being used and $S\simeq 0.8$ is a safety factor.\\
If the error is larger than the user-specified tolerance, the time step is rejected. A new time step size is estimated using  \eqt{eq:edt2} as well.

To sum up, each solution undergoes Picard iterations until convergence before the error is calculated; for IQS P-C, each solution is re-scaled by its corresponding amplitude $p$.  Additionally, if the nonlinear iteration does not converge, the entire step doubling process is repeated with half the time original time step.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RATTLESNAKE Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

RATTLESNAKE is a MOOSE-based application developed INL specific to solving radiation transport problems with multiphysics capabilities.  MOOSE is a finite-element based, multiphysics framework that gives the general architecture for the development of physics application like RATTLESNAKE.   At the heart of RATTLESNAKE is the action system, which provides a means to consolidate the MOOSE input syntax (which can be quite larger for multigroup transport simulations), so that a user does not have to define every kernel, variable, etc., used in the problem. Rather, the user inputs an equation description 
(e.g., Diffusion, S$_n$, P$_n$, etc.) and a solution method (e.g., SAAF, LS, CFEM, DFEM, etc.), and the action system will incorporate all the necessary physics involved (kernels, boundary conditions, postprocessor, etc.).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Executioner}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

An IQS excecutioner was created to implement the IQS method. The IQS executioner derives from the Transient executioner in MOOSE.  The IQS executioner contains a loop over micro time steps that computes the PRKE solution and then passes $p$ and $\frac{dp}{dt}$ for the Transient executioner to evaluate the shape equation at each macro step.  The PRKE solve is performed with a user specified option of backward-Euler, Crank-Nicholson, or SDIRK33.   The IQS executioner also supplements Transient Picard iteration process by adding its own error criteria:
\be
Error_{IQS}=\left|\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,n}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^{g,0}\right)}-1\right|
\label{eq:eiqs}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Action System}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The IQS implementation mostly require the specific IQS executioner, described above. However, IQS additional changes are in the RATTLESNAKE action system in order to support the IQS execution.  First, changes needed to be made in order to evaluate the shape equation.  The shape equation, after some manipulation, is very similar to the time-dependent governing laws that RATTLESNAKE already solves. Using multigroup diffusion as an example again, we show the various kernels employed and highlight the new or modified kernels.

\begin{align}
\frac{\partial}{\partial t}\left(\frac{\varphi^g}{v^g}\right)=&\underbrace{\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'}}_{Flux Kernel} + \underbrace{\sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'}}_{Flux Kernel} - \underbrace{\left( -\div D^g \grad \right)\varphi^g}_{Flux Kernel} - \underbrace{\Sigma_r^g\varphi^g}_{Flux Kernel} \nonumber \\
& - \underbrace{\frac{1}{v^g} \boxed{\overbrace{\frac{1}{p}\frac{dp}{dt}}^{From Executioner}}\varphi^g}_{IQS Kernel}+\underbrace{\frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i}_{Modified Flux Kernel}
\end{align}

To enable RATTLESNAKE to solve this equation, an IQS removal kernel was created to evaluate $\sum_{g=1}^G\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\varphi^g$ and added when the IQS executioner is called.  Also, the precursor kernel was modified to include the $\frac{1}{p}$ term. Finally, the precursor auxkernel that evaluates \eqt{eq:preq2} using the analytical integration method described in \sct{sect:prec}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PRKE coefficients}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to evaluate the PRKE coefficients, defined by Equations~\ref{eq:rmb}~-~\ref{eq:l}, four postprocessors were created.  The parameter calculations were separated by $\frac{\bar{\beta}_i}{\Lambda}$ numerator, $\bar{\lambda}_i$ numerator/denominator, $\frac{\rho-\bar{\beta}}{\Lambda}/\frac{\bar{\beta}}{\Lambda}$ denominator, and $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator. The first three are relatively simple, only relying on material properties and solution quantities, then computing the elemental integral.  The $\frac{\rho-\bar{\beta}}{\Lambda}$ numerator requires the use of the MOOSE \texttt{save\_in} feature. This feature saves the residual from a calculated kernel or boundary contribution in the shape evaluation to an auxiliary variable.  The postprocessor then computes the inner product of this variable and the initial adjoint solution.  After each of these postprocessors are evaluated, a user object pulls together all the values and performs the numerator/denominator divisions.  The resulting values are then passed to the executioner for the PRKE solve.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Other Action Systems}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For simplicity, IQS implementation has only been described for CFEM diffusion.  However, RATTLESNAKE has other action systems capable of transient simulation, where IQS can be implemented and be effective.  One of these action systems is DFEM diffusion, where the only major difference from CFEM is the diffusion term in Equations~\ref{eq:flux}~and~\ref{eq:shape2}.  However, in the derivation for IQS, this term is unaffected between shape and flux evaluation.  So saving the residual for this diffusion kernel in the \texttt{save\_in} variable is the only alteration to this action for IQS to function.\\

Additional action systems involve transport, but it is evident from \sct{sect:transport} that IQS implementation in these is straightforward as well.  The main differences between a diffusion implementation and a transport implementation are outlined below:
\begin{enumerate}
\item The form of the operators in the shape equations is different, but RATTLESNAKE has already implemented all the kernels necessary to represent these operators.  So no change is necessary to these kernels is necessary for IQS.  Additionally, the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ is the same, so both action systems can use the same kernel.
\item The PRKE parameters also change because of the operators.  For the $(\rho-\bar{\beta})$ parameter, the same post-processor can be used with the \texttt{save\_in} functionality in MOOSE.  The post-processor for $\bar{\beta}_i$ must be re-written for transport, but takes a very similar form as diffusion.
\end{enumerate}

IQS will be made available to any action system capable of transient simulation, which includes:
\begin{itemize}
\item CFEM-Diffusion
\item DFEM-Diffusion
\item SAAF-CFEM-SN
\item SAAF-CFEM-PN
\item LS-CFEM-SN
\item DFEM-SN
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Predictor-Corrector Modification}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In order to preserve the already implemented standard version of IQS, an option in the IQS executioner was created to specify which method is desired.  Because the diffusion solve is flux instead of shape, when predictor-corrector option is specified, the IQS removal kernel ($\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$) and the modified precursor kernel are bypassed, while all the postprocessors are still executed.  However, it is difficult to rescale the flux to shape before the PRKE parameter postprocessor are executed.  So the parameters are computed using the full flux, but amplitude is space independent and comes out of the integrals.  As seen in parameter definitions, when shape is replaced with flux, the amplitude comes out of the integral and cancels out.  So the conversion of the predicted flux in  \eqt{eq:rescale} to shape is unnecessary if the corrected flux is solved with \eqt{eq:phicorr2}.  After obtaining the corrected flux, the precursors are re-evaluated using a \texttt{EXEC\_LINEAR} statement.
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}} p_{n+1}
\label{eq:phicorr2}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Input}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The input file for IQS is very similar to the current transient diffusion input file.  The IQS input has a different executioner type and parameters.  The executioner type is simply IQS and input parameters include the number of micro time steps per macro step, the IQS error tolerance, and the initial power.  The RATTLESNAKE transient action system currently requires a multi-app and transfer to compute and pass the initial $\phi$ and $\keff$, which is present in the transient input deck.  However, IQS also requires an initial evaluation of the adjoint flux, for the weighting function.  So another input file and multi-app transfer was made for the adjoint calculation.  \\

Below is the syntax for the executioner block for an IQS input file.  The \texttt{predictor\_corrector} logical determines whether to do IQS P-C or regular IQS.  The \texttt{IQS\_error\_tol} is the tolerance for the IQS error represented by \eqt{eq:eiqs} and will at most \texttt{picard\_max\_its} iterations until convergence.  The \texttt{prke\_scheme} defines the time discretization used for the PRKE solution. Where \texttt{RK} is SDIRK33, \texttt{CN} is Crank-Nicholson, and \texttt{IE} is implicit-Euler.
\begin{lstlisting}
[Executioner]
  type = IQS
  predictor_corrector = true/false
  picard_max_its = 5
  ...
  n_micro = 10000
  IQS_error_tol = 1e-7
  prke_scheme = 'RK'
[]
\end{lstlisting}

Since IQS needs to use the adjoint solution for the PRKE parameter evaluation, auxiliary variables need to be created for each group in the input file.  Below is an example of their definition: 
\begin{lstlisting}
[AuxVariables]
  [./adjoint_flux_g0] 
    family = LAGRANGE
    order = FIRST
  [../]
  [./adjoint_flux_g1]
    family = LAGRANGE
    order = FIRST
  [../]
  ...
[]
\end{lstlisting}

Below is the syntax for the multi-app block to perform forward and adjoint steady-state evaluations:
\begin{lstlisting}
[MultiApps]
 [./initial_solve]
   type = FullSolveMultiApp
   execute_on = initial
   input_files = initial.i
 [../]
 [./adjoint_solve]
   type = FullSolveMultiApp
   execute_on = initial
   input_files = adjoint.i
 [../]
[]
\end{lstlisting}

Below is the syntax for the Transfer block to copy the initial and adjoint solutions from the multi-apps.
\begin{lstlisting}
[Transfers]
 [./copy_solution]
   type = TransportSystemVariableTransfer
   direction = from_multiapp
   multi_app = initial_solve
   execute_on = initial
   from_transport_system = diff
   to_transport_system = diff
 [../]
 [./copy_adjoint]
   type = MultiAppVariableTransfer
   execute_on = initial
   direction = from_multiapp
   multi_app = adjoint_solve
   from_variables = 'sflux_g0 sflux_g1 ...'
   to_variables = 'adjoint_flux_g0 adjoint_flux_g1 ...'
 [../]
 [./copy_eigenvalue]
   type = EigenvalueTransfer
   execute_on = initial
   direction = from_multiapp
   multi_app = initial_solve
 [../]
[]
\end{lstlisting}

\newpage
%------------------------------------------------------------------------------
%
%------------------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{RESULTS} 
\label{sect::results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This section describes results of examples that test the IQS implementation and shows its effectiveness on computation speed and accuracy.  Three examples were selected for this purpose.  The first is a homogeneous one-group problem, subjected to a heterogeneous material change (absorption cross-section change as a ramp in time for a subset of the geometry).  The second is the two-dimensional TWIGL ramp transient benchmark, described further.  The final is the LRA benchmark, which a two-dimensional, temperature feedback example.  Each example was solved with regular diffusion (brute force), IQS, and IQS P-C.  Each method was tested with backward-Euler and BDF2 for the first two examples; just Crank-Nicholson was used for third example.  The performance for each method are represented by convergence plots with constant $\Delta t$ and by the number of time steps with step doubling time adaptation.  For time adaptation, backward-Euler time discretization was used.
% and, for a safety factor, $e_{tol} = 0.8e_{max}$. I found tol and max to be confusing to the reader. i have simplified a bit

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{One-Dimensional Custom Example}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The example is very simple and computes quickly, it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.  \fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.  Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.  The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.

\begin{figure}[!htbp]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}
\caption{1-D slab region identification}
\label{fig:slab}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab material properties and problem parameters}
\label{tab:1Dmat}
\begin{tabular}{llllll}
\hline
$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
\hline
1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\

\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab absorption cross-section at times of interest}
\label{tab:1Dslope}
\begin{tabular}{llllllll}
\hline
Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
\end{center}
\end{table}

\fig{fig:power} shows the power at each macro time step as compared to the traditional brute force (full flux time discretization) method.  The strong correlation between the two curves shows that IQS is consistent with a proven method for a highly transient example.  \fig{fig:conv} shows that IQS is not only consistent for this example, but also has a better error constant in the convergence study.  \fig{fig:conv} shows the error convergence for brute force and IQS methods.  This plot shows that IQS has consistent convergence slope for first and second order time schemes and that IQS has a better error constant for the convergence.  \tbl{tab:1Ddt2} shows the results for time adaptation.  These results show that traditional IQS does not perform well with step doubling time adaptation and IQS P-C performs moderately better than brute force.  The IQS performance in this example are not exceptional because it is not an easy problem.  As seen in \fig{eq:D_shape}, the shape undergoes significant change through the transient, which requires many shape evaluations to retain accuracy.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.49\textwidth]{\FiguresDir/1D_het_power_plot.png}
\includegraphics[width=0.5\textwidth]{\FiguresDir/1D_het_power_plot2.png}
\caption{Power level of 1D example}
\label{fig:power}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=5in]{\FiguresDir/1D_het_convergence.jpg}
\caption{Error convergence comparison of 1D example}
\label{fig:conv}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab step doubling results}
\label{tab:1Ddt2}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05 	&	0.021596 &	10  &	32	 &	0.10084 &	12	  &	175		&	0.028019 &	10	&	33 \\
2 &	0.01 	&	0.032864 &	21  &	80	 &	0.08030 &	41	  &	850	 	&	0.044721 &	20	&	74 \\
3 &	0.005 	&	0.053159 &	27  &	96	 &	0.01215 &	115	  &	2220 	&	0.052095 &	25	&	96 \\
4 & 0.001 	&	0.056546 &	56  &	188	 &	0.03242 &	848	  &	12511 	&	0.061274 &	50	&	162 \\
5 &	0.0005 	&	0.062882 &	77  &	245	 &	0.03491 &	1718  &	25841 	&	0.061240 &	69	&	225 \\
6 &	0.0001 	&	0.060089 &	177 &	537	 &	0.03554 &	8702  &	129985 	&	0.060824 &	159	&	480 \\
7 &	5.0e-05	&	0.059513 &	252 &	767	 &	0.03622 &	17282 &	256463 	&	0.061078 &	224	&	680 \\
8 &	1.0e-05	&	0.061063 &	561 &	1691 &	0.03142 &	79988 &	1104227 &	0.061901 &	501	&	1509 \\
\hline
\end{tabular}}
\end{center}
\end{table}

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=\linewidth]{\FiguresDir/1D_shape.png}
\caption{Flux profile of 1D slab at various times}
\label{fig:1D_shape}
\end{center}
\end{figure}

\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{TWIGL Benchmark}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.  It is a 2D, 2-group reactor core model with no reflector region shown in \fig{fig:TWIGL_reg}.  \tbl{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
\caption{TWIGL benchmark problem description}
\label{fig:TWIGL_reg}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{TWIGL benchmark material properties and slope perturbation}
\label{tab:TWIGL_mat}
\begin{tabular}{llllllll}
\hline
  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
\hline
1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
\hline
  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
\hline
  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
\hline
 \multicolumn{8}{l}{\footnotesize Material 1 ramp perturbation:} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
\end{tabular}
\end{center}
\end{table}

\fig{fig:TWIGL_power} shows the IQS  solution as compared with the Brute Force solution.  These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake.  Finally, \fig{fig:TWIGL_conv} plots the error convergence of IQS and the Brute Force methods.  The curves show the impressive convergence of IQS for the highly transient TWIGL example.  \tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for time adaptation.  The results show that both IQS methods perform exceptionally well compared to brute force.  It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.

%It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot because the amplitude term is not included, but the gradients of colors is comparable.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.49\linewidth]{\FiguresDir/TWIGL_power_plot.png}
\includegraphics[width=0.5\linewidth]{\FiguresDir/TWIGL_power_plot2.png}
\caption{Power level comparison of TWIGL Benchmark}
\label{fig:TWIGL_power}
\end{figure}

%\begin{figure}[!htbp]
%\begin{center}
%\begin{subfigure}[!htbp]{0.4\textwidth}
%\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
%\caption{Brute force flux}
%\end{subfigure}
%\quad
%\begin{subfigure}[!htbp]{0.4\textwidth}
%\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
%\caption{IQS Shape}
%\end{subfigure}
%\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
%\label{fig:TWIGL_plots}
%\end{center}
%\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=5in]{\FiguresDir/TWIGL_convergence.jpg}
\caption{Error convergence comparison of TWIGL Benchmark}
\label{fig:TWIGL_conv}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{TWIGL step doubling results}
\label{tab:TWIGLdt2}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	7   &	65   &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
\hline

\end{tabular}}
\end{center}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\linewidth]{\FiguresDir/TWIGL_power_plot_dt2.png}
\caption{Power level comparison of TWIGL Benchmark with time adaptation}
\label{fig:TWIGL_power_dt2}
\end{figure}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LRA Benchmark}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient.  To have better understanding on the cross sections given later, we present the equations here:
\begin{subequations}
\begin{align}
-\frac{1}{v_1} \frac{\partial \phi_1}{\partial t} &= -\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 - \nu(1-\beta)S_f  - \sum_{i=1}^2 \lambda_i C_i, \\
-\frac{1}{v_2} \frac{\partial \phi_2}{\partial t} &= -\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 - \Sigma_{s, 1\rightarrow 2}\phi_1, \\
S_f &= \sum_{g=1}^2 \Sigma_{f,g} \phi_g, \\
\frac{\partial C_i}{\partial t} &= \nu\beta_i f - \lambda_i C_i, \quad i=1,2, \\
\frac{\partial T}{\partial t} &= \alpha f, \label{eq:lra-temp} \\
\Sigma_{a,1} &= \Sigma_{a,1}(\vec{r}, t=0) \left[1+\gamma\left(\sqrt{T} - \sqrt{T_0}\right)\right], \\
P &= \kappa S_f,
\end{align}
\end{subequations}
where $\phi_1$, $\phi_2$ are the fast and thermal fluxes; $v_1, v_2$ are the averaged neutron velocities; $\Sigma_{a,1}, \Sigma_{a,2}$ are the absorption cross sections; $\Sigma_{s,1\rightarrow 2}$ is the fast-to-thermal scattering cross section; $\Sigma_{f,1}, \Sigma_{f,2}$ are the fission cross sections; $\nu$ is the averaged number of neutrons emitted per fission; $\beta_1, \beta_2$ are the delayed neutron precursor fractions and $\beta=\beta_1 + \beta_2$; $C_1, C_2$ are the delayed neutron precursor concentrations; $\lambda_1, \lambda_2$ are the decay constants of the delayed neutron precursors; $S_f$ is the fission reaction rate; $P$ is the power density; $T$ is the temperature; $\kappa$ is the averaged power released per fission; $\alpha$ is the combination of $\kappa$ and the specific heat capacity; $\gamma$ is the Doppler feedback coefficient; $T_0=T(\vec{r}, t=0)$.
The two-group diffusion equation are solved with zero flux boundary conditions on external surfaces, reflecting conditions at symmetry boundaries and steady state initial conditions which are obtained by solving
\begin{align}
-\div D_1 \grad\phi_1 + (\Sigma_{a,1} + \Sigma_{s, 1\rightarrow 2})\phi_1 &= \frac{1}{k}\sum_{g=1}^2 \nu\Sigma_{f,g}\phi_g, \\
-\div D_2 \grad\phi_1 + \Sigma_{a,2}\phi_2 =& \Sigma_{s, 1\rightarrow 2}\phi_1. \\
\end{align}
The eigenvalue $k$ is used to modify the fission cross section for the transient simulations with $\frac{1}{k}\Sigma_{f,g}, g=1,2$.  The initial flux distribution shall be normalized such that the averaged power density
\begin{align}
\bar{P} \equiv \frac{\int_{V_{core}} P(\vec{r}, t=0) d\vec{r}}{\int_{V_{core}} d\vec{r}},
\end{align}
where $V_{core}$ is the core region with fuels, is equal to $10^{-6} W\cdot cm^{-3}$.
The initial precursor concentrations are in equilibrium with the initial critical flux distribution.\\

The geometry is illustrated in~\fig{fig:lra-geometry}.\\
\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\linewidth]{\FiguresDir/lra.png}
\caption{LRA benchmark geometry with region assignment.}
\label{fig:lra-geometry}
\end{figure}

Initial two-group constants are presented in~\tbl{tab:lra-xs}.
$\nu$ is equal to 2.43.
Axial bulking $B^2 = 10^{-4}$ is applied for both energy groups.
Delayed neutron data are presented in~\tbl{tab:lra-dnp}.
All fuel materials have the same delayed neutron data.
Some scalar data are listed in~\tbl{tab:lra-scalar}.\\
\begin{table}[!htbp]
    \centering
    \caption{LRA benchmark initial two-group constants.\label{tab:lra-xs}}
    \resizebox{\textwidth}{!}{
      \begin{tabular}{|c|c|c|c|l|l|l|l|l|}
      \hline
             &                 & Group   & $D_g$      & $\Sigma_{a,g}$ & $\nu\Sigma_{f,g}$ & $\Sigma_{s,1\rightarrow 2}$ & $\chi_g$ & $v_g$              \\
      Region & Material        &       g & (cm)       &  ($cm^{-1}$)   &  ($cm^{-1}$)      &  ($cm^{-1}$)                &          & ($cm\cdot s^{-1}$) \\
      \hline
      1      & Fuel 1 with rod & 1       & 1.255      & 0.008252       & 0.004602          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.211      & 0.1003         & 0.1091            & 0.02533                     & 0        & $3.0\times10^5$    \\
      \hline
      2      & Fuel 1 without rod & 1    & 1.268      & 0.007181       & 0.004609          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.1902     & 0.07047        & 0.08675           & 0.02767                     & 0        & $3.0\times10^5$    \\
      \hline
      3      & Fuel 2 with rod & 1       & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                 & 2       & 0.2091     & 0.08344        & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      4      & Fuel 2 without rod & 1    & 1.259      & 0.008002       & 0.004663          &                             & 1        & $3.0\times10^7$    \\
             &                    & 2    & 0.2091     & 0.073324       & 0.1021            & 0.02617                     & 0        & $3.0\times10^5$    \\
      \hline
      5      & Reflector        & 1      & 1.257      & 0.0006034      & -                 &                             & -        & $3.0\times10^7$    \\
             &                  & 2      & 0.1592     & 0.01911        & -                 & 0.04754                     & -        & $3.0\times10^5$    \\
      \hline
      \end{tabular}}
\end{table}

\begin{table}[!htbp]
    \centering
    \caption{LRA benchmark delayed neutron data.\label{tab:lra-dnp}}
      \begin{tabular}{|c|l|l|l|l|}
      \hline
      Group i & $\beta_i$ & $\lambda_i$ ($s^{-1}$) & $\chi_{d,i,1}$ & $\chi_{d,i,2}$ \\
      \hline
      1       & 0.0054    & 0.0654  & 1 & 0 \\
      2       & 0.001087  & 1.35    & 1 & 0 \\
      \hline
      \end{tabular}
\end{table}

\begin{table}[!htbp]
    \centering
    \caption{LRA benchmark scalar values.\label{tab:lra-scalar}}
      \begin{tabular}{|l|c|l|}
      \hline
      Meaning & Notation & value \\
      \hline
      Axial buckling for both energy groups & $B_g^2$  & $10^{-4}$ ($cm^{-2}$)\\
      Mean number of neutrons per fission   & $\nu$    & 2.43 \\
      Conversion factor                     & $\alpha$ & $3.83\times 10^{-11}$ ($K\cdot cm^{3}$) \\
      Feedback constant                     & $\gamma$ & $3.034\times 10^{-3}$ ($K^{1/2}$) \\
      Energy released per fission           & $\kappa$ & $3.204\times 10^{-11} $ ($J/fission$) \\
      Initial and reference temperature     & $T_0$    & 300 (K) \\
      Active core volume                    & $V_{core}$ & 17550 ($cm^2$)\\
      \hline
      \end{tabular}
\end{table}

The transient is initiated by changing the thermal absorption cross section as the following:
\begin{align}
\Sigma_{a,2}(t) = \Sigma_{a,2}(t=0) \left\{\begin{array}{lr} 1-0.0606184t, & t\leq 2 \\
                                                             0.8787631, & t>2
                                           \end{array}\right.
\end{align}
where $t$ is time in seconds.\\

The \fig{fig:LRA_plots} plots show the IQS  solution as compared with the Brute Force solution.    These plots show that IQS is consistent with Doppler feedback problems in RATTLESNAKE.  \fig{fig:lra_conv} plots the error convergence of IQS and the Brute Force methods. \tbl{tab:LRAdt2} shows the time adaptation performance of brute force and IQS P-C.  These results show that IQS, again, performs much better than the brute force method by taking significantly fewer diffusion solves while retaining similar accuracy.

\begin{figure}[!htbp]
\begin{center}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile.png}
\caption{Power Profile}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_power_profile2.png}
\caption{Power Profile at Peak}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/lra_temp_profile.png}
\caption{Temperature Profile}
\end{subfigure}
\caption{LRA Benchmark power and temperature comparison}
\label{fig:LRA_plots}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.6\linewidth]{\FiguresDir/lra_convergence.jpg}
\caption{Error convergence comparison of LRA Benchmark}
\label{fig:lra_conv}
\end{figure}

\begin{table}[!htbp]
\caption{LRA step doubling results}
\label{tab:LRAdt2}
\begin{center}
\begin{tabular}{|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	 0.01 &	 0.0437566 &	 145 &	 735 &	 0.143855 &	 105 &	 616 \\ 
2 &	 0.005 &	 0.0309732 &	 266 &	 948 &	 0.104782 &	 157 &	 814 \\ 
3 &	 0.001 &	 0.158242 &	 585 &	 2046 &	 0.155676 &	 218 &	 1062 \\ 
4 &	 0.0005 &	 0.270949 &	 833 &	 2908 &	 0.211994 &	 480 &	 2038 \\ 
5 &	 0.0001 &	 0.280468 &	 1793 &	 6155 &	 0.0437761 &	 434 &	 1458 \\ 
6 &	 5e-05 &	 0.212274 &	 2485 &	 8511 &	 0.0385312 &	 542 &	 1719 \\ 
\hline
\end{tabular}
\end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

The purpose of this report is to briefly review the improved quasi-static method for neutronic transient simulations, to describe its implementation in RATTLESNAKE, the radiation transport application based on MOOSE, and to show its effectiveness using several test cases.  
%The derivation of IQS was done with multigroup diffusion, but it was seen in \sct{sect:transport} that application to transport is relatively simple.  
Two IQS implementations are available in RATTLESNAKE.  The first one is the traditional IQS approach, where the shape equation and PRKE are solved as a coupled nonlinear system with Picard iterations.  The second approach, IQS P-C, is a predictor-corrector linearization, where the flux is solved and then rescaled using the amplitude obtained through the PRKE
(Point Reactor Kinetic Equations).  IQS P-C is a simpler method and is computationally much less expensive to perform. \\

IQS's implementation in RATTLESNAKE involves a new kernel, an auxiliary kernel, a new executioner derived from the current transient, four postprocessors, as well as alterations to the action system.  The new kernel implements the $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term in the shape equation.  The auxiliary kernel alters how the precursors are solved, the method is outlined by \sct{sect:prec}.  The executioner is where the PRKE are evaluated after shape solves are performed.  The postprocessors are used to evaluate the parameters for the PRKE and a user object is used to pass them to the executioner.  IQS's implementation the action system made sure that all the correct kernels and postprocessors are executed when the IQS executioner is specified in the input. \\

The IQS implementations have been verified and comparing against the traditional ``brute force'' temporal discretization of the governing laws. Error convergence plots and time adaptation performance are provided.  Three examples were used including a custom one-dimensional example, the TWIGL benchmark, and LRA benchmark.  The one-dimensional example shows large shape changes in the flux and proves consistency of the shape evaluation. IQS shows proper error convergence for first and second order time discretization. IQS P-C shows marginal improvement with time adaptation, but traditional IQS shows poor performance.  The TWIGL benchmark shows the effectiveness of the PRKE in IQS because there is very little shape change through the transient.  IQS shows very satisfactory performance with both constant time steps and time adaptation.  The LRA benchmark tests IQS when utilized in the context of a multiphysics simulation.  Again, IQS shows good performance by requiring half the time steps as brute force for a given error.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\clearpage
\pagebreak
\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrvnat}
\bibliography{references_IQS}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Appendix: aaaa}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\tcr{putting stuff here for now}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%It can be shown that the most appropriate weighting function $(w^g)$ is the initial adjoint flux $(\phi^{*g})$.  For brevity, the following definition will be applied: $\int_D\phi^{*g}(\vec{r})f(\vec{r})d^3r=\left(\phi^{*g},f\right)$
%\begin{align}
%& \quad \frac{dp}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)+ p\frac{d}{dt}\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)= \nonumber \\
%& p\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)  \\
%& + \sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i) \nonumber
%\end{align}
%\be
%\frac{d}{dt}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)=\frac{1}{\keff}\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})p - \sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \quad 1 \le i \le I 
%\ee
%
%In order to impose uniqueness of the factorization, one imposes:
%\be
%\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)=constant
%\ee
%
%Therefore the PRKE formulation reduces to:
%\begin{align}
%\frac{dp}{dt}= &\frac{\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta) \sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g}  \varphi^{g'} - \left( -\div D^g \grad  + \Sigma_r^g \right) \varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p \nonumber \\ 
%& +\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i
%\end{align}
%\be
%\frac{d\xi_i}{dt}=\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}p - \frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}\xi_i \quad 1 \le i \le I 
%\ee
%Where:
%\be
%\xi_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g C_i)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
%\ee
%
%It is convenient to define the effective reactivity, delay-neutron fraction, and delayed-neutron precursor decay constant:
%\be
%\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff}(1-\beta)\sum_{g'=1}^G \nu^{g'} \Sigma_f^{g'}\varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -\left( -\div D^g \grad  + \Sigma_r^g \right)\varphi^g\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
%%\label{eq:rmb}
%\ee
%\be
%\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{1}{\keff}\frac{\sum_{g=1}^G(\phi^{*g}, \chi_{d,i}^g\beta_i\sum_{g'=1}^G\nu^{g'} \Sigma_f^{g' }\varphi^{g'})}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
%%\label{eq:b}
%\ee
%\be
%\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
%%\label{eq:l}
%\ee
%
%So:
%\be
%\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
%%\label{eq:p}
%\ee
%\be
%\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
%%\label{eq:c}
%\ee
%
%Equations (\ref{eq:p}) and (\ref{eq:c}) are a formulation of the point reactor kinetics equation (PRKE), but the parameters (Equations (\ref{eq:rmb})-(\ref{eq:l})) are dependent on the shape.  If the assumption is made that the shape is time-independent, the shape is computed once at the first time step and used for the PRKE parameter evaluation at all other steps.  However, if the shape is dependent on time, the shape needs to be computed in transient using equation (\ref{eq:shape}) and (\ref{eq:prec}) in order retain accuracy.  Equations (\ref{eq:shape2}) and (\ref{eq:preq2}) shows the usual form of the shape and precursor equations with amplitude put on the right hand side.  Equation (\ref{eq:shape2}) is very similar to the multigroup flux equation (\ref{eq:flux}), except the removal cross-section term is augmented by a $\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}$ term and the precursor contribution has a $\frac{1}{p}$ multiplier.  Equation (\ref{eq:preq2}) is very similar to the normal precursor equation (\ref{eq:precursor}), except the fission source term is multiplied by $p$.  These differences are crucial for IQS implementation in Rattlesnake.
%
%\begin{align}
%\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t} = &\frac{\chi_p^g}{\keff} (1-\beta)\sum_{g'=1}^G  \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} \nonumber \\ 
%& -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g + \frac{1}{p}\sum_{i=1}^I\chi_{d,i}^g\lambda_iC_i  , \quad 1 \le g \le G 
%%\label{eq:shape2}
%\end{align}
%\be
%\frac{dC_i}{dt} = \frac{\beta_i}{\keff}p \sum_{g=1}^G\nu^{g} \Sigma_f^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
%%\label{eq:preq2}
%\ee

%\begin{align}
%& \quad \sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\frac{\partial\varphi^g}{\partial t}\right)= \nonumber \\ 
%&\sum_{g=1}^G\left(\phi^{*g},\frac{\chi_p^g}{\keff} \sum_{g'=1}^G (1-\beta) \nu^{g'} \Sigma_f^{g'} \varphi^{g'} + \sum_{g'\neq g}^G\Sigma_s^{g'\to g} \varphi^{g'} -  \left( -\div D^g \grad  + \Sigma_r^g + \frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}\right) \varphi^g\right) \\
%&+\frac{1}{p}\sum_{i=1}^I\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_iC_i) \nonumber
%\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}