### 2017-02-04 doc/Thesis/data/section3.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson
%	 Version 3.16.09
%  Last updated 9/12/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION III
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{KINETICS EXAMPLES}

Neutron dynamics is the study of the time-dependent nature of neutrons in a reactor.  More specifically, there is no coupling of neutron behavior with other physics, like thermal hydraulic feedback.  This section describes kinetics examples that IQS is tested with and analysis of its performance.  The examples range in complexity and application.  The first is a one-dimensional problem, designed for the prototype code in MATLAB.  The next four are from the Argonne National Lab (ANL) Benchmark Problem Book (BPB), and are common problems for testing codes and developing methods \cite{ANL_BPB}.

\section{One Dimensional Prototype Problem}

This example is very simple and computes quickly; it entails a one dimensional, homogeneous 400 cm slab with a heterogenous perturbation in absorption cross section.  \fig{fig:slab} shows how the regions of the slab are divided and \tbl{tab:1Dmat} shows the initial material properties.  Regions 2, 3, and 4 have slope perturbations at different points in time, \tbl{tab:1Dslope} shows the values of the absorption cross-section in each region at the times of interest.  The values of $\Sigma_a$ between these times of interest are linear interpolations between the given values.

\begin{figure}[!htbp]
\begin{center}
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}
\caption{1-D slab region identification}
\label{fig:slab}
\end{center}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab material properties and problem parameters}
\label{tab:1Dmat}
\begin{tabular}{llllll}
\hline
$D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu \Sigma_f (cm^{-1})$ & $v (cm/s)$ & $\beta$ & $\lambda (s^{-1})$ \\
\hline
1.0 & 1.1 & 1.1 & 1,000 & 0.006 & 0.1 \\

\hline
\end{tabular}
\end{center}
\end{table}

\begin{table}[!htbp]
\begin{center}
\caption{1-D slab absorption cross-section at times of interest}
\label{tab:1Dslope}
\begin{tabular}{llllllll}
\hline
Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
\end{center}
\end{table}

\subsection{IQS Iteration Convergence}

\subsection{Time Step Convergence}

\pagebreak
\section{TWIGL Benchmark}

This benchmark problem originates from the Argonne National Lab Benchmark Problem Book \cite{ANL_BPB}.  It is a 2D, 2-group reactor core model with no reflector region shown in \fig{fig:TWIGL_reg} \cite{TWIGL_benchmark}.  This example is meant to be of progressive complexity from the previous example.  The transient of this reactor is very geometrically symmetrical with very little temporal shape change.  Therefore, IQS is expected to perform significantly better than the implicit discretization method.  \tbl{tab:TWIGL_mat} shows the material properties of each fuel region and the ramp perturbation of Material 1.

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.75\textwidth]{\FiguresDir/TWIGL_regions.jpg}
\caption{TWIGL benchmark problem description}
\label{fig:TWIGL_reg}
\end{center}
\end{figure}
\begin{table}[!htbp]
\begin{center}
\caption{TWIGL benchmark material properties and slope perturbation}
\label{tab:TWIGL_mat}
\begin{tabular}{llllllll}
\hline
  &  &  &  &  &  &  \multicolumn{2}{c}{$\underline{\Sigma_s (cm^{-1})} $} \\
Material & Group & $D (cm)$ & $\Sigma_a (cm^{-1})$ & $\nu\Sigma_f (cm^{-1})$ & $\chi$ & $g \rightarrow 1$ & $g \rightarrow 2$ \\
\hline
1 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01 \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
2 & 1 & 1.4 & 0.010 & 0.007 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.4 & 0.150 & 0.200 & 0.0 & 0.0 & 0.00  \\
3 & 1 & 1.3 & 0.008 & 0.003 & 1.0 & 0.0 & 0.01  \\
  & 2 & 0.5 & 0.050 & 0.060 & 0.0 & 0.0 & 0.00  \\
\hline
  & $\nu$ & $v_1 (cm/s)$ & $v_2 (cm/s)$ & $\beta$ & $\lambda (1/s)$ &   &   \\
\hline
  & 2.43 & 1.0E7 & 2.0E5 & 0.0075 & 0.08 &   &   \\
\hline
 \multicolumn{8}{l}{\footnotesize Material 1 ramp perturbation:} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (1-0.11667t) \quad t \leq 0.2 s$} \\
\multicolumn{8}{l}{\footnotesize $\Sigma_{a,2}(t)=\Sigma_{a,2}(0) \times (0.97666t) \quad t > 0.2 s$} \normalsize
\end{tabular}
\end{center}
\end{table}

\subsection{TWIGL Convergence Analysis}

Figs. \ref{fig:TWIGL_power} and \ref{fig:TWIGL_plots} show the IQS  solution as compared with the implicit discretization solution.  It is important to note the IQS shape plot is scaled differently than the Brute Force flux plot (\fig{fig:TWIGL_plots}) because the amplitude term is not included, but the gradients of colors is comparable. These plots show that IQS is consistent in more complex, higher dimensional problems in Rattlesnake. These plots also serve to illustrate that IQS has a much more accurate solution, even at a significantly larger time step than the implicit discretization. In order to demonstrate asymptotic convergence of IQS, implicit Euler (IE) and second order BDF (BDF2) were applied to the TWIGL simulation. \fig{fig:TWIGL_conv} plots the error convergence of IQS and the implicit discretization methods.  The curves show the impressive convergence of IQS for the highly transient TWIGL example. The slope indicated in the legend are the linear slope of curves on the log plot, these slopes should be similar to the order of the method (1 for IE and 2 for BDF2).  IQS shows a increased order because the PRKE is performing much of accuracy convergence and it is computed using SDIRK33, a third order method.

\begin{figure}[!htbp]
\centering
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/TWIGL_power_plot.png}
\caption{Power profile for entire transient}
\end{subfigure}
\begin{subfigure}[!htbp]{0.49\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/TWIGL_power_plot2.png}
\caption{Power at cusp of profile}
\end{subfigure}
\caption{Power level comparison of TWIGL Benchmark}
\label{fig:TWIGL_power}
\end{figure}

\begin{figure}[!htbp]
\begin{center}
\begin{subfigure}[!htbp]{0.4\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/ndiff_ramp_flux.jpg}
\caption{Implicit Discretization flux}
\end{subfigure}
\quad
\begin{subfigure}[!htbp]{0.4\textwidth}
\includegraphics[width=\textwidth]{\FiguresDir/iqs_ramp_shape.jpg}
\caption{IQS Shape}
\end{subfigure}
\caption{TWIGL Benchmark flux/shape comparison at $t=0.2$}
\label{fig:TWIGL_plots}
\end{center}
\end{figure}

\begin{figure}[!htbp]
\centering
\includegraphics[height=3in]{\FiguresDir/TWIGL_convergence.jpg}
\caption{Error convergence comparison of TWIGL Benchmark}
\label{fig:TWIGL_conv}
\end{figure}

\subsection{TWIGL with Step Doubling Time Adaptation}

\tbl{tab:TWIGLdt2} and \fig{fig:TWIGL_power_dt2} show the results for TWIGL with time adaptation.  The results show that both IQS methods perform exceptionally well compared to implicit discretization.  It also shows that traditional IQS performed better with large $e_{tol}$, while IQS P-C was better with smaller $e_{tol}$.

\begin{table}[!htbp]
\begin{center}
\caption{TWIGL step doubling results}
\label{tab:TWIGLdt2}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|l|l|l|l|}
\hline
  &  & \multicolumn{3}{|c|}{Implicit Discretization} & \multicolumn{3}{|c|}{IQS} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves & Error & Steps & Solves & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9   &	29  &	0.03380433 &	4   &	20   &	0.00323100 &	4  &	9   \\
2 &	0.01    &	3.5555e-05 &	11  &	35  &	0.00166991 &	5   &	40   &	0.00263068 &	5  &	12  \\
3 &	0.005   &	4.0364e-05 &	11  &	31  &	0.00886584 &	5   &	40   &	0.00160486 &	6  &	21  \\
4 &	0.001   &	0.00294822 &	33  &	122 &	0.02976305 &	5   &	36   &	1.7527e-05 &	10 &	35  \\
5 &	0.0005  &	0.00099778 &	39  &	131 &	0.00143781 &	6   &	55   &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	0.00019510 &	78  &	236 &	0.00016175 &	8   &	65   &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	0.00018372 &	112 &	342 &	6.0328e-05 &	12  &	163  &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05 &	8.0564e-05 &	263 &	794 &	7.7103e-05 &	379 &	5729 &	9.8321e-07 &	48 &	210 \\
\hline

\end{tabular}}
\end{center}
\end{table}

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.8\linewidth]{\FiguresDir/TWIGL_power_plot_dt2.png}
\caption{Power level comparison of TWIGL Benchmark with time adaptation}
\label{fig:TWIGL_power_dt2}
\end{figure}

\pagebreak
\section{C5G7-TD Benchmark}

The C5G7 benchmark is a MOX fueled, pressurized water reactor (PWR) minicore configuration. The C5G7 geometry depicted in Fig. \ref{fig:c5g7-2D} comprises a 2-by-2 array of UO$_2$ and MOX assemblies surrounded by moderator. In Fig. \ref{fig:c5g7-2D} vacuum and reflective boundary conditions are denoted by V and R, respectively. The fuel pins are not homogenized, each pin cell is comprised of a fuel pin, fission chamber, or guide tube surrounded by moderator as depicted in Fig. \ref{fig:c5g7-pincell}. The cladding and gap are homogenized into the fuel and are not explicitly modeled. Each assembly is made up of a regular 17-by-17 grid of pin cells.
Seven energy-group cross sections for the seven material regions are given in \cite{c5g7}. The energy group boundaries are provided in Table \ref{tab:c5g7-ebounds}; three of the seven energy groups are fast, four are thermal.

\begin{figure}[!htpb]
\begin{center}
\includegraphics[width=\linewidth]{\FiguresDir/C5G7-2D.png}
\end{center}
\caption{Geometry of the two-dimensional C5G7 benchmark problem. Boundary conditions are vacuum (V) or reflective (R).}
\label{fig:c5g7-2D}
\end{figure}

\begin{figure}[!htpb]
\begin{center}
\includegraphics[width=0.5\linewidth]{\FiguresDir/C5G7-pincell.png}
\end{center}
\caption{Geometry of a C5G7 pin-cell.}
\label{fig:c5g7-pincell}
\end{figure}

\begin{table}[!htpb]
\centering
\caption{Energy group boundaries for C5G7 MOX benchmark.}
\label{tab:c5g7-ebounds}
\begin{tabular}{c c}
\hline
Group & Upper Energy\\
\hline
 1&20 MeV  \\
 2&1 MeV  \\
 3&500 keV  \\
 4&3 eV  \\
 5&0.625 MeV  \\
 6&0.1 MeV  \\
71&0.02 MeV  \\
\hline
\end{tabular}
\end{table}

\pagebreak
\section{LMW Benchmark?}