### 2017-02-16 doc/M&C_2017/FullPaper/MC2017_FullPaper.tex
Adding more results

\documentclass{anstrans}

%%%% packages and definitions (optional)
\usepackage{color} % allows inclusion of graphics
\usepackage{graphicx} % allows inclusion of graphics
\usepackage{booktabs} % nice rules (thick lines) for tables
\usepackage{microtype} % improves typography for PDF
\usepackage{subcaption}
\usepackage{tikz}

\newcommand{\SN}{S$_N$}
\renewcommand{\vec}[1]{\bm{#1}} %vector is bold italic
\newcommand{\vd}{\bm{\cdot}} % slightly bold vector dot
\newcommand{\ud}{\mathop{}\!\mathrm{d}} % upright derivative symbol
%  new definitions
\newcommand{\bs}[1]{\mathbf{#1}}
\renewcommand{\div}{\bs{\nabla}\! \cdot \!}
\newcommand{\grad}{\bs{\nabla}}
% extra space
\newcommand{\qq}{\quad\quad}
% common reference commands
\newcommand{\eqt}[1]{Eq.~(\ref{#1})}                     % equation
\newcommand{\fig}[1]{Fig.~\ref{#1}}                      % figure
\newcommand{\tbl}[1]{Table~\ref{#1}}                     % table
\newcommand{\sct}[1]{Section~\ref{#1}}                   % section
\newcommand{\app}[1]{Appendix~\ref{#1}}                   % appendix

\newcommand{\keff}{k_\textit{eff}}

\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\vn}{\vec{n}}
\newcommand{\vel}{\vec{\mathrm{v}}}
\newcommand{\adj}{\Phi^\dagger_0}
\newcommand{\tcr}[1]{\textcolor{red}{#1}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert_{L^2}}

%%%% changes from the original 'anstrans' class
\usepackage{fancyhdr} % allows headers and footers

\renewcommand\headrule{} % remove underline in the header
\setcounter{secnumdepth}{1}
\renewcommand{\thesection}{\Roman{section}.}
\renewcommand{\thesubsection}{\arabic{subsection}.}
\renewcommand{\thesubsubsection}{\Alph{subsubsection}.}
\makeatletter
\renewcommand*{\@seccntformat}[1]{\csname the#1\endcsname\hspace{1mm}}
\makeatother

%%%% Header
\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{\fontsize{9}{9} \itshape
M\&C 2017 - International Conference on Mathematics \& Computational Methods Applied to Nuclear Science \& Engineering,
\\ Jeju, Korea, April 16-20, 2017, on USB (2017)
}

%%%% Maketitle
\title{Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method}
\author{Zachary M.~Prince, Jean C.~Ragusa}

\institute{Department of Nuclear Engineering, Texas A\&M University, College Station, TX}

\email{zachmprince@tamu.edu \and jean.ragusa@tamu.edu}

% tikz stuff
\usetikzlibrary{shapes,arrows,positioning}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}
\tikzstyle{greenblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=green!20, ultra thick, minimum width=60mm, minimum height=15mm,]
\tikzstyle{blueblock}=[rectangle, rounded corners, draw, align=center, top color=white, bottom color=blue!20, ultra thick, minimum width=60mm, minimum height=15mm]
\tikzstyle{reddiamond}=[diamond, draw, align=center, top color=white, bottom color=red!20, ultra thick, minimum width=60mm, aspect=2]

%%%% Abstract
\begin{document}
\vspace*{-42pt}
\begin{strip}
\centering{\parbox{153mm}{{\bf Abstract} \itshape - 
The purpose of this paper is to present a new technique in dealing with multi-physics feedback in transient reactor simulations, specifically utilizing the improved quasi-static method (IQS).  the technique involves the addition of an intermediate time scale to evaluate temperature and PRKE parameters.  The theory is that temperature and the parameters are more quickly varying in time than the shape, thus further reducing the number of multi-group diffusion evaluations and execution time for a given accuracy. The technique was implemented into the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) and tested with the LRA benchmark and the Transient-15 TREAT problem.  The results proved that the addition of this time scale was effective in reducing execution time for these problems while maintaining accuracy. 
}\par}
\vspace*{14pt}
\end{strip}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The purpose of this paper is to introduce several techniques in dealing with multi-physics feedback for transient neutron diffusion calculations.  In dynamic simulations, the neutron temporal distribution in a nuclear core can be strongly influenced by non-neutronics variables, e.g., temperature. The improved quasi-static method (IQS) is an effective technique for simulating the temporal behavior of the neutron flux in reactors. Here, we present a study combining the IQS method with multi-physics solution techniques for coupled transient calculations.

IQS is a transient spatial kinetics method that involves factorizing the neutron flux into a space- and time-dependent component (shape) and a time-dependent component (amplitude) \cite{Ott_1966,Dulla2008}. The technique relies on the shape being less rapidly varying in time compared to the flux, hence requiring fewer shape computations or updates. IQS has largely been applied to neutron kinetics, without other feedback mechanisms. This paper presents the application of multi-physics feedback with IQS and analyzes performance with temperature feedback benchmark problems.

The majority of publicized applications of IQS involve purely neutron kinetics, with several exceptions.  Devooght et. al. discusses the application of the generalized quasi-static method with thermal hydraulic feedback within a newton iteration scheme in \cite{Devooght_1984}. This implementation is not ideally efficient for IQS because, theoretically, the shape is less variant in time than temperature feedback and temperature is more closely coupled with amplitude.  Other references such as \cite{Banfield_2013} and \cite{KIKO3D_2003} apply IQS to multi-physics feedback problems, but are ambiguous on their coupling treatment.

This paper presents a new approach to multi-physics feedback with IQS involving inclusion of feedback mechanisms in the quasi-static process.  The intention of this implementation is to further optimize solution accuracy with computational effort.  In order to evaluate the performance of the methodology, it will be tested with two temperature feedback problems: the LRA benchmark and an example from Transient Reactor Testing Facility (TREAT) reactor.  The performance is quantified by comparing accuracies with traditional implementations of neutron dynamics and IQS.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In this Section, we recall the equation for the IQS method, starting from the standard multi-group diffusion equations with delayed neutron precursors in operator form:

\be
\frac{1}{v^g}\frac{\partial \phi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \phi^{g'} - L^g\phi^g + S_{d}^g
\label{eq:flux}
\ee 
\be
\frac{dC_i}{dt} = \sum_{g=1}^G P_{d,i}^g \phi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:precursor}
\ee

Where $H^{g'\to g}$ is the scattering operator, $P_p^{g' \to g}$ is the prompt fission operator, $L^g$ is the streaming and interaction operator, 
$S_{d}^g$ is the delay source, and $P_{d,i}^g$ is the delayed-neutron fission operator.

The flux factorization approach leads to a decomposition of the multigroup flux into the product of a time-dependent amplitude ($p$) and a space-/time-dependent 
multigroup shape ($\varphi$):
\be
\phi^g(\vec{r},t)=p(t)\varphi^g(\vec{r},t)
\ee
After implementing the factorization, the shape diffusion equations result:
\be
\frac{1}{v^g}\frac{\partial \varphi^g}{\partial t} = \sum_{g'=1}^G \left(H^{g'\to g} + P_p^{g' \to g} \right) \varphi^{g'} - \left(L^g - \boxed{\frac{1}{v^g}\frac{1}{p}\frac{dp}{dt}}\right)\varphi^g + \boxed{\frac{1}{p}}S_{d}^g
\label{eq:shape}
\ee 
\be
\frac{dC_i}{dt} = \boxed{p}\sum_{g=1}^G P_{d,i}^g \varphi^{g} - \lambda_i C_i \ , \quad 1 \le i \le I 
\label{eq:prec}
\ee

Note that the time-dependent shape equation is similar to the time-dependent flux equation. 
However, the shape/amplitude equations are now non-linearly coupled (boxed terms).

To obtain the amplitude equations, the multigroup shape equations are multiplied by a weighting function, typically the initial adjoint flux ($\phi^{*g}$), and then integrated over the domain.  For brevity, the inner product over space will be represented with parenthetical notation ($\left(\phi^{*g},f^g\right) = \int_D \phi^{*g}(\vec{r})f^g(\vec{r})d^3r
$). In order to impose uniqueness of the factorization, one requires $\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)$ to be constant.  And after some manipulation, the standard point reactor kinetics equations (PRKE) for the amplitude solution are obtained:
\be
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\ee
\be
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}p-\bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\ee
Where the functional coefficients are calculated using the space-/time-dependent shape function as follows:
\be
\frac{\rho-\bar{\beta}}{\Lambda}=\frac{ \sum_{g=1}^G\left(\phi^{*g},\sum_{g'}(H^{g' \to g}+P_p^{g' \to g}-L^{g'}\delta_{g'g})\varphi^{g'}\right)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\label{eq:rmb}
\ee
\be
\frac{\bar{\beta}}{\Lambda}=\sum_{i=1}^I\frac{\bar{\beta}_i}{\Lambda}=\sum_{i=1}^I\frac{\sum_{g=1}^G(\phi^{*g}, P_{d,i}^g \varphi^g)}{\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g\right)}
\ee
\be
\bar{\lambda}_i=\frac{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^g\lambda_i C_i)}{\sum_{g=1}^G(\phi^{*g},\chi_{d,i}^gC_i)}
\label{eq:l}
\ee

Solving for the shape in \eqt{eq:shape} can become expensive, especially in two or three dimensions, and even more so when using the transport equations in lieu of the diffusion equations.  Using IQS, one expects the time dependence of the shape to be weaker than that of the flux itself,  thus allowing for larger time step sizes in updating the shape. The PRKE equations form a small ODE system and can be solved using a much smaller time step size. In transients where the shape varies much less than the flux, IQS can be very computationally effective. The two-time scale solution process, a micro scale for the PRKE and a macro scale for the shape, is illustrated in \fig{fig:iqsviz}.

\begin{figure}[!htbp]
\centering
\includegraphics[width=\linewidth]{IQS_visualization.jpg}
\caption{IQS method visualization}
\label{fig:iqsviz}
\end{figure}

It is important to note that the PRKE parameters are evaluated at each macro step and linearly interpolated for the PRKE evaluation. In order to preserve the error convergence of high order discretization schemes for shape, higher order interpolation of the parameters is required.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS Predictor-Corrector (IQS P-C)}

To avoid performing the IQS non-linear solve, a predictor-corrector version of IQS has been derived in \cite{Dulla2008}. One first solves the neutron flux (represented by Equations \ref{eq:flux} and \ref{eq:precursor}) to obtain a predicted flux. The predicted flux is then converted to a shape through a rescaling argument :
\be
\varphi^g_{n+1} = \underbrace{\phi^g_{n+1}}_{\text{predicted}} \frac{K_0}{K_{n+1}}
\label{eq:rescale}
\ee
Where:
\be
K_{n+1} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{n+1}\right)
\ee
\be
K_{0} =\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\varphi^g_{n+1}\right)=\sum_{g=1}^G\left(\phi^{*g},\frac{1}{v^g}\phi^g_{0}\right)
\ee

The PRKE parameters are then computed with this shape using Equations \eqref{eq:rmb}-\eqref{eq:l} and interpolated over the macro step for the solution of the PRKE equations.  With the newly computed amplitude, the shape is rescaled and the corrected flux is evaluated:
\be
\underbrace{\phi^g_{n+1}}_{\text{corrected}} = p_{n+1} \times \varphi^g_{n+1} \,.
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{IQS Solution Process with Multiphysics}

Other physical quantities, such as temperature, are affected by reaction rates and subsequently affect the operators of the flux equations.  
For IQS, this feedback affects both the shape equation and the parameters of the PRKE; thus, it is an additional nonlinear component to the 
already non-linear shape-amplitude equations.  Each of these components have different temporal dependencies; so it may be beneficial for 
efficiency to evaluate them on different time scales.  The amplitude is more rapidly varying than the shape which is computationally expensive to evaluate
and is evaluated only on macro-time steps. In multiphysics simulations, one may take advantage of a fine-scale power distribution in the coupled
physics components (temperature). Figure~\ref{fig:time} shows a such a solution process for temperature feedback.

\begin{figure}[htbp!]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[scale=1]
\tikzset{font={\footnotesize}}
%Shape
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(shape) at (-.25,2) {$\varphi:$};

%Temp/Params
\draw[] (0,1) -- (6,1) ;
\foreach \x in  {0,6}
\draw[shift={(\x,1)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {2,4}
\draw[shift={(\x,1)},color=black] (0pt,4pt) -- (0pt,-4pt);
\draw[shift={(0,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,1)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(temp) at (-.75,1) {$T \rightarrow \rho$:};

% PRKE
\draw[] (0,0) -- (6,0) ;
\foreach \x in  {0,6}
\draw[shift={(\x,0)},color=black] (0pt,6pt) -- (0pt,-6pt);
\foreach \x in  {0,2,4,6}
\draw[shift={(\x,0)},color=black] (0pt,4pt) -- (0pt,-4pt);
\foreach \x in  {0,0.2,0.4,0.6,0.8,1,1.2,1.4,1.6,1.8,2,2.2,2.4,2.6,2.8,3,3.2,3.4,3.6,3.8,4,4.2,4.4,4.6,4.8,5,5.2,5.4,5.6,5.8,6}
\draw[shift={(\x,0)},color=black] (0pt,2pt) -- (0pt,-2pt);
\draw[shift={(0,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(6,0)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\node(prke) at (-.5,0) {$p, \frac{dp}{dt}:$};

\draw (0,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 2} (2,0);
\draw (2,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 4} (4,0);
\draw (4,0) edge[out=45,in=135,->,thick] node[above,sloped] {Step 6} (6,0);
\draw (0,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 1} (2,1);
\draw (2,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 3} (4,1);
\draw (4,1) edge[out=45,in=135,->,thick] node[above,sloped] {Step 5} (6,1);
\draw (0,2) edge[out=15,in=165,->,thick] node[above,sloped] {Step 7} (6,2);

\draw (0,2) edge[out=-135,in=135,->,thick] node[below,sloped] {$\varphi$} (0,1);
\draw (2,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (2,0);
\draw (2,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (2,1);
\draw (4,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (4,0);
\draw (4,0) edge[out=45,in=-45,->,thick] node[below,sloped] {$p$} (4,1);
\draw (6,1) edge[out=-135,in=135,->,thick] node[below,sloped] {$\rho$} (6,0);
\draw (6,0) edge[out=45,in=0,->,thick] node[below,sloped] {$p, \frac{dp}{dt}$} (6,2);

\end{tikzpicture}}
\caption{Time scales and process of IQS}
\label{fig:time}
\end{figure}

The top time scale represents a shape diffusion evaluation on a macro step, the middle has an arbitrary three steps within the macro step where temperature and the PRKE parameters are evaluated, and the bottom one represents the PRKE evaluation on micro steps.  The shape is linearly interpolated within the macro step for the temperature and PRKE parameter evaluation, and the parameters are interpolated within the temperature step for the PRKE evaluation.  Since there is a nonlinear coupling between all these components, each temperature step is iterated until amplitude has converged and the macro step is iterated until shape has converged. \fig{fig:proc} further visualizes this process as a programming structure.

\begin{figure}[!htpb]
\centering
\resizebox{\linewidth}{!}{%
\begin{tikzpicture}[every node/.style = {font=\normalsize}]

\node[greenblock](p1) at (0,0) {Save Old Solution};
\node[greenblock](p2) at (0,-2.5) {Save Current Solution};
\node[greenblock](p3) at (0,-5) {Interpolate Solution};
\node[blueblock](p4) at (0,-7.5) {Solve Temperature};
\node[greenblock] (p5) at (0,-10) {Update PRKE Params};
\node[greenblock] (p6) at (0,-12.5) {Solve PRKE};
\node[reddiamond] (check1) at (0,-17.5) {Check for \\ convergence ($p$)};
\node[reddiamond] (check2) at (7,-12) {If $t=t_{n+1}$};
\node[greenblock] (p7) at (8,-6) {Shape Solve};
\node[reddiamond] (check3) at (8,-2.5) {Check for \\ convergence ($\varphi$)};
%\node [above =0.1mm of p1] {\large{IQS Solve:}};
%\node [above =0.1mm of p5] {\large{Multi-physics:}};

%\tikzback{p1}{p1}{p4}{p4}{bk1}
%\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- node[right] {$t=t_{n+1}$}(p2.north);
\draw[->,ultra thick](p2.south) -- node[right] {$t=t_{n}+\Delta t/3$}(p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)-- (p5.north);
\draw[->,ultra thick](p5.south) -- (p6.north);
\draw[->,ultra thick](p6.south) -- (check1.north);
\draw[->,ultra thick](check1.west) -- node[above,sloped] {\large{no}} (-5,-17.5) |-  (p4.west);
\draw[->,ultra thick](check1.east) -| node[right] {\large{yes}} (check2.south);
\draw[->,ultra thick](check2.west) -- node[above,sloped] {\large{no}\small \qq $t=t+\Delta t/3$} (4,-5) -- (p3.east);
\draw[->,ultra thick](check2.north) --  (8,-10) -- node[right] {\large{yes}} (p7.south);
\draw[->,ultra thick](p7.north) -- (check3.south);
\draw[->,ultra thick](check3.west) -- node[above] {no} (p2.east);
\draw[->,ultra thick](check3.north) -- node[right] {yes} (8,0) -- node[above] {$t_n=t_{n+1}$} (p1.east);

\end{tikzpicture}}
\caption{Visualization of shape iteration and temperature update process for IQS}
\label{fig:proc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Dynamical Time Scale Analysis}
\label{sect:tau}

The temporal variance of flux, shape, temperature, and amplitude can be quantified by defining a dynamical time scale ($\tau$) for each physics.  A small value $\tau$ means the variable is quickly varying in time and consequently needs relatively small time steps for accuracy, vice-versa for large $\tau$.  The general definition of $tau$ is defined by \eqt{eq:tau1}, where $\theta$ is the variable of interest.  Since each variable in discretized in time, a finite difference approximation will be made for the $\frac{d\theta}{dt}$ term and the average between the two corresponding time steps will be made for the $\frac{1}{\theta}$ term.  Additionally, $\tau$ is spatially dependent for flux and temperature,  but only the time dependent behaivior of this quantity is of interest.  Therefore, the $L^2$ norm of each term will be used to compute the approximate time scale ($\tilde{\tau}$), formally defined by \eqt{eq:tau2}.  $\theta$ represents a summation over groups for flux and shape.

\be
\tau = \frac{1}{\left|\frac{1}{\theta}\frac{d\theta}{dt}\right|}
\label{eq:tau1}
\ee

\be
\tilde{\tau}_{n+1} = \frac{\norm{\theta_{n+1} + \theta_{n}}/2}{\norm{\theta_{n+1} - \theta_{n}}/\Delta t}
\label{eq:tau2}
\ee

According to the a priori hypothesis from previous sections, $\tau$ is large for shape, somewhat smaller for temperature, and very small for amplitude and flux.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Analysis}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
.
To test the multiple time scale implementation of IQS, the LRA benchmark was chosen for reference and a TREAT example was employed.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{LRA Benchmark}

The LRA benchmark is a two-dimensional, two-group neutron diffusion problem with adiabatic heat-up and Doppler feedback in thermal reactor \cite{ANL_BPB}.  It is a super prompt-critical transient. The execution of the benchmark was performed by the Rattlesnake/MOOSE framework at Idaho National Laboratory (INL) \cite{wang2013}.  The spacial discretization was performed using continuous finite element method with first order Lagrangian basis functions. The mesh consisted of blocks 11X11 with five uniform refinements, totaling 165,165 elements and 124,609 nodes. Three different temporal techniques were applied: implicit discretization of the flux equation (Brute force), IQS, and IQS-PC. Crank-Nicholson time discretization scheme was used for the diffusion evaluation of each technique.  Third order Runge-Kutta discretization with step doubling adaptation was used for the PRKE evaluation.  The performance of IQS and the temperature updates were measured by its improvement in accuracy at peak power over the Brute force technique.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{LRA Temperature Feedback}
\label{sec:LRA_T}

The heat up is described by \eqt{eq:temp} and the feedback is described by \eqt{eq:dopp}.

\be
\rho c_p \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
\label{eq:temp}
\ee

\be
\Sigma_a^{thermal}(\vec{r},t) = \Sigma_a^{thermal}(\vec{r},0)\left[1+\gamma\left(\sqrt{T}-\sqrt{T_0}\right)\right]
\label{eq:dopp}
\ee

In the temperature evaluation, a typical implicit solver would simply use the interpolated flux at end of the temperature time step for the right-hand-side of the equation.  However, IQS has much more information about the profile of the flux along the time step because of the micro-step amplitude evaluation.  Therefore, it is possible to solve for temperature using a semi-analytical approach, shown by \eqt{eq:temp_an}.

\be
T^{n+1} = T^n + \frac{\kappa_f}{\rho c_p} \left(a_2 \varphi^{n+1} + a_1 \varphi^{n}\right)
\label{eq:temp_an}
\ee
Where $n$ corresponds to the beginning of the temperature step.  $a_1$ and $a_2$ are integration coefficients defined by \eqt{eq:a1} and \eqt{eq:a2}.
\be
a_1 = \int_{t_n}^{t_{n+1}}\left(\frac{t_{n+1}-t'}{\Delta t}\right)p(t')dt'
\label{eq:a1}
\ee
\be
a_2 = \int_{t_n}^{t_{n+1}}\left(\frac{t'-t_n}{\Delta t}\right)p(t')dt'
\label{eq:a2}
\ee

Any interpolation of the amplitude function $p(t)$ along the micro steps is possible to carry out the above macro-step integrals. Here, we use piece-wise linear interpolation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{LRA Results}

\fig{fig:lra_profile} shows the baseline power and temperature profile for the LRA benchmark.  The baseline results are compared to the results achieved by Sutton and Aviles in \cite{Sutton_1996} and presented in \tbl{tab:base}.  The relative difference in the magnitude of the peak power ($t\approx1.44 s$) from the baseline was used for error comparison.  \fig{fig:lra_bad} is an error convergence plot comparing the three techniques where temperature is evaluated only on the macro step (1 temperature update).  \fig{fig:conv} is an error convergence plot comparing the three techniques where temperature is evaluated 5 times within a macro step (5 temperature updates).  Finally, \fig{fig:mp} shows the effect of various temperature updates. The dashed lines correspond to brute force at different flux step sizes, while the IQS macro step size is kept constant. \\

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{lra_profile.png}
\caption{LRA baseline temperature and power profile}
\label{fig:lra_profile}
\end{figure}

\begin{table}[!htbp]
\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{|l|cc|}
\hline
Calculation  &  Baseline & Sutton (Spandex 1936) \\
\hline
No. of Spatial Nodes	& 3872 		& 1936 \\
Eigenvalue 				& 0.99637	& 0.99637 \\
No. of Time Steps 		& 6000 		& 23,890 \\
Time to Peak Power (s) 	& 1.441 	& 1.441 \\
Peak Power (W/cm$^3$) 	& 5456 		& 5461 \\
\hline
\end{tabular}
}
\end{center}
\caption{LRA baseline verification}
\label{tab:base}
\end{table}

\begin{figure}[!htpb]
\centering
\includegraphics[width=\linewidth]{lra_bad.jpg}
\caption{LRA convergence plots with only one temperature update each macro step}
\label{fig:lra_bad}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{lra_mp_convergence.png}
\caption{Error convergence plot with 5 temperature updates per macro step}
\label{fig:conv}
\end{figure}

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{lra_mp.png}
\caption{Error plot with various temperature updates per macro step}
\label{fig:mp}
\end{figure}

The convergence plots show that updating temperature and the PRKE parameters within a macro step has a significant effect on the performance of IQS.  With only one update, IQS was only slightly better than brute force, brute force required about 150\% more time steps than IQS for the same error.  While 5 temperature updates showed a much more significant IQS performance, brute force required about 400\% more time steps than IQS for the same error.  \fig{fig:mp} shows that error has a convergent behavior for the number of temperature updates.  This convergence makes sense because temperature can only be so accurate before the error in shape is dominating. Table \ref{tab:ndiff_lra} shows the run time results for the brute force calculations.  Tables \ref{tab:iqs_lra} and \ref{tab:iqspc_lra} present the IQS run-times with various numbers of temperature updates.  These run-times are based on total alive time of the execution where the diffusion evaluation is distributed over 24 processors. These run-times show a marginal performance for IQS and impressive performance for IQS P-C.  Some of the execution times were able to decrease from brute force with the same number of macro steps because IQS is better equipped to resolve the nonlinearity between temperature and amplitude. Furthermore, there does seem to be an ideal number of temperature updates to optimize execution time: IQS only needs one and IQS P-C seems to be ideal at 4 updates. This discrepancy in the number of updates shows that a adaptive type implementation of the updates would be ideal, and could enforce a constant error over the transient. It is also important to compare the error of brute force with IQS at one update and IQS P-C at 4 updates.  IQS shows an error comparable to brute force at $\Delta t = 0.002$, signifying an actual increase in runtime by -34.1\%.  IQS P-C shows an error less than brute force at $\Delta t = 0.002$, signifying an actual increase in runtime by <-34.9\%. \\

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
Run  &  $\Delta t$ & Error & Runtime (hr) & Linear Iter.\\
\hline
1	& 4.0e-3	& 1.407e-2 	& 4.11	& 7.13e4	\\
2	& 2.0e-3	& 3.174e-3 	& 6.01	& 9.49e4 	\\
3 	& 1.0e-3 	& 7.690e-4 	& 10.38	& 1.45e5	\\
4 	& 5.0e-4 	& 1.892e-4 	& 21.91	& 2.08e5	\\
5 	& 2.5e-4	& 4.590e-5 	& 25.23	& 3.16e5	\\
\hline
\end{tabular}
%}
\end{center}
\caption{Brute force run time results}
\label{tab:ndiff_lra}
\end{table}

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 2.612e-3 	& 3.96 	& -3.18\%	\\
2	& 2		& 9.893e-4 	& 6.02	&  47.1\%	\\
3 	& 4 	& 5.796e-4 	& 7.87	&  92.3\%	\\
4 	& 8 	& 4.772e-4 	& 12.61	& 207.9\% 	\\
5 	& 16	& 4.516e-4 	& 22.14	& 440.7\%	\\
\hline
\end{tabular}
%}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from $\Delta t = 0.004$ brute force 
\caption{IQS run time results with $\Delta t = 0.004$}
\label{tab:iqs_lra}
\end{table}

\begin{table}[!htbp]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|l|ccc|}
\hline
	&  Temperature 	&  		& Runtime 	& \% Increase	\\
Run	&  Updates 	& Error & (hr)		& in Runtime$^*$\\
\hline
1	& 1		& 3.488e-3 	& 2.91 	& -28.9\%	\\
2	& 2		& 1.349e-3 	& 3.73	& -9.00\%	\\
3 	& 4 	& 9.161e-4 	& 3.97	& -3.04\%	\\
4 	& 8 	& 8.052e-4 	& 5.39	&  31.7\%	\\
5 	& 16	& 7.905e-4 	& 8.19	&  100\%	\\
\hline
\end{tabular}
%}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from $\Delta t = 0.004$ brute force 
\caption{IQS PC run time results with $\Delta t = 0.004$}
\label{tab:iqspc_lra}
\end{table}

The performance of the quasi-statics can also be explained by the computation of the dynamical time scale described by \sct{sect:tau}.  \fig{fig:LRAtc} shows the time scale profile over the transient, computed using \eqt{eq:tau2}.  This plot shows that in a brute force simulation, the flux dominates the time dependent behaivior, while temperature lags in its variance for the majority of the transient.  In an IQS simulation, the time scale behavior of amplitude almost exactly matches the flux, while shape is more varying than temperature throughout most of the transient.  The large $\tau$ for temperature during the beginning of the transient shows that adaptation of the number of updates is important; computational expense on temperature evaluations is being wasted during this time.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{time_constant_lra.png}
\caption{Dynamical time scale for LRA benchmark}
\label{fig:LRAtc}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transient-15 TREAT Example}

Transient 15 is a test case based on the TREAT core. The purpose of the original creation of this simulation in Rattlesnake is to test the model's fidelity with the thermal feedback of TREAT, but it is not meant to exactly match any previous experiments.  Nevertheless, the goal of the following simulations is to test IQS and its time scale based treatment of temperature with a more complex model. Transient 15 involves an 11-energy group diffusion approximation and is discretized into $355,712$ hexahedral continuous finite elements totaling $4,109,523$ degrees of freedom.  The three-second transient involves a linear ramp decrease in the absorption cross section throughout the control rod region. \fig{fig:Tran15} shows a visualization of the flux profile within the core, hidden is the massive amount of graphite surrounding the core.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{Tran15_core2.png}
\caption{Transient 15 core power profile at peak power}
\label{fig:Tran15}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Transient-15 Temperature Feedback}

The Transient-15 model uses a adiabatic temperature feedback mechanism, similar to the one explored by the LRA. \eqt{eq:temp2} describes the heat up of the fuel.  It is very similar, except the specific heat is now dependent on temperature is described by \eqt{eq:cp}.  The temperature evaluation is identical to the one described in LRA section, except a Newton iteration process is employed to resolve the nonlinearity from the specific heat term.  The feedback to the cross-sections are applied using linear interpolation of tabular data provided by INL.

\be
\rho c_p(T) \frac{\partial T(\vec{r},t)}{\partial t} = \kappa_f \sum^G_{g=1}\Sigma_f^g \phi^g(\vec{r},t)
\label{eq:temp2}
\ee

\be
c_p = -5.8219e-10T^3 - 4.3694e-7T^2 + 2.8369e-3T -1.009e-2
\label{eq:cp}
\ee

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Transient-15 Results}

In order to test the temperature feedback treatment, six different scenarios were run: a baseline with a very small time step, brute force, IQS with one and 5 temperature updates per macro step, and IQS P-C with one and 5 updates.  \fig{fig:Tran15_profile} shows the baseline power and temperature profile for the Transient-15 example.  \tbl{tab:tran15} shows the error and runtime results.

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{Tran15_profile.png}
\caption{Transient 15 total power and average temperature profile during transient}
\label{fig:Tran15_profile}
\end{figure}

\begin{table*}[htb]
\begin{center}
%\resizebox{\linewidth}{!}{
\begin{tabular}{|l|ccccccc|}
\hline
 & No.	& Max &	Time at Max &	Max Average &	\% Increase &	Max Power & Linear	\\
Method & of Steps &	Power (W) &	Power (s) &	Temperature (K) & Runtime$^*$ & Error & Iterations \\
\hline
Baseline 			& 3000 	& 3.5039e+08 & 1.901 & 371 &	---		&	---		 & ---		\\
Brute Force 		& 300 	& 3.5011e+08 & 1.90  & 371 &	---		&	7.875e-4 & 41020	\\
IQS 				& 300	& 3.5036e+08 & 1.90  & 371 & -11.9\%	&	8.385e-5 & 23949	\\
IQS (5 updates) 	& 300 	& 3.5040e+08 & 1.90  & 371 &  49.7\%	&	3.687e-5 & 24035	\\
IQS P-C 			& 300 	& 3.5065e+08 & 1.90  & 371 &  -2.1\%	&	7.527e-4 & 39020	\\
IQS P-C (5 updates) & 300 	& 3.5043e+08 & 1.90  & 371 &  26.5\%	&	1.227e-4 & 37866	\\
\hline
\end{tabular}
%}
\end{center}
\vspace{-3mm}
$^*$ difference in runtime from brute force 
\caption{Transient-15 Error and Runtime Results}
\label{tab:tran15}
\end{table*}

The results from \tbl{tab:tran15} show similar performance of IQS with the temperature updates as the LRA.  IQS with 1 temperature update shows a performance that reduces the error to approximately a tenth of the brute force error, and reduces the execution time by about 12\%.  This shows that IQS was able to resolve the nonlinearity between flux and temperature with significantly less diffusion evaluations.  Having IQS with 5 updates significantly increased the execution time for the same time step, but the error was reduced.  Comparing this error to a similar brute force error at a smaller time step could show that the runtime was reduced. IQS P-C performed not nearly as well as it did with the LRA benchmark, but still proved to be effective.  Having 5 updates for IQS P-C increased the runtime marginally, but decreased the error significantly.  The transient profile of the varibles' dynamical time scales is shown in \fig{fig:tran15tc}.  This plot exhibits a similar response to that of the LRA. The response of temperature shows that the updates are a computational frugal treatment of the feedback and adaptation of the number of updates is vital for optimization. \\

\begin{figure}[htbp!]
\centering
\includegraphics[width=\linewidth]{time_constant_tran15.png}
\caption{Dynamical time scale for the Transient 15 example}
\label{fig:tran15tc}
\end{figure}

%\newpage
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

IQS is a powerful tool for reactor simulation, it has the capability of greatly reducing computation time while retaining accuracy of relevant quantities.  The goal of this paper is to demonstrate IQS's performance with multi-physics simulations, as well as techniques to improve its performance.  The technique focused on incorporating a new time scale in IQS for feedback quantities and PRKE parameters. The methodology is that for problems involving prompt-critical transients, temperature and its feedback to amplitude is more quickly varying than shape, but less time-dependent than amplitude itself.  Therefore, an intermediate time scale was implemented, as well as a fixed-point iteration between amplitude and temperature.  The attempt is to reduce the number of multi-group diffusion evaluations in a transient by reducing the number of time steps required for a given accuracy, and reduce the number of nonlinear iterations between flux/shape and temperature.  Since multi-group diffusion evaluations are significantly more computationally expensive than temperature and amplitude evaluations, this result will show a decrease in execution time for a given accuracy.

Two examples were chosen to test the implementation of the temperature feedback treatment: the LRA benchmark and the Transient-15 TREAT problem.  Both problems involve adiabatic heat up of the core, the LRA has a constant specific heat, while Transient-15's is dependent on temperature.  The LRA showed clear results concluding a marginal improvement in performance for IQS and a impressive performance for IQS P-C.  The results proved that the updates improved the performance for both IQS and IQS P-C, but optimization of the number of updates is important, possibly requiring adaptation.  The implementation to the Transient-15 problem gave similar results.  IQS proved to marginally benefit from the updates, decreasing error but increasing execution time.  IQS P-C was benefited more by the updates in this example. The updates were able to reduce the error with a marginal increase in execution time.

In conclusion, the multi-physics treatment with IQS presented in this paper is worth further investigation, optimization, and implementation into other transient reactor problems with more feedback mechanisms.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\appendix
%\section{Appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
This work was supported by the Department of Energy, Idaho National Laboratory, and the Integrated University Program Fellowship.  We thank Mark Dehart, Yaqi Wang, the NEAMS program, and INL's Moose/Rattlesnake team for their support.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{ans}
\bibliography{references_IQS}

\end{document}