### 2017-02-23 doc/Thesis/data/introduction.tex
Thesis: Addressing initial comments

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  New template code for TAMU Theses and Dissertations starting Fall 2016.
%
%  Author: Sean Zachary Roberson 
%	 Version 3.08.16
%  Last updated 8/19/2016
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                           SECTION I
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\pagestyle{plain} % No headers, just page numbers
\pagenumbering{arabic} % Arabic numerals
\setcounter{page}{1}

\chapter{\uppercase{Introduction}}

\section{Background on Transient Reactor Testing}

The primary purpose of transient reactor testing is the safety and performance analysis of fuel and other reactor components. Since the March 2011 events at the Fukushima Daiichi Nuclear Power Plant in Japan, there has been a significant governmental demand for the development of accident tolerant fuels in light water reactors (LWR) \cite{ConRepTREAT}.  Currently, two facilities are capable of testing fuels under these extreme conditions: Annular Core Research Reactor (ACRR) at Sandia National Laboratory (SNL) and the Transient Reactor Testing Facitlity (TREAT) at Idaho National Laboratory (INL) \cite{SandiaTREAT}. TREAT had been put in stand-by status in 1994, but in 2014 a Final Environmental Assessment by the Department of Energy (DOE) approved the restart of the facility \cite{FONSI}.

TREAT is an graphite-moderated, air-cooled, thermal reactor built and operated starting in 1959.  The reactor was designed to subject fuels and experimental apparatus to extreme power pulses, for fast reactor type components \cite{TREATSummary}. The primary purpose for the restart of the facility is to use these pulses to simulate accident scenarios and test the fidelity of accident tolerant LWR fuels. TREAT is expected to resume experimentation and testing by 2020.  In addition to the substantial upgrades of TREAT's electronic and mechanical systems during the restart, advance computer models of the reactor are in development.  These models have a mutualistic relationship with TREAT: validating fuel models with experimentation and using reactor models to streamline experimental procedures. Development of these models have renewed a significant interest in implementing transient reactor simulation methods in modern computational tools.

Transient reactor simulation methods have been developed for several decades. The revitalization of TREAT and other transient reactor testing efforts has brought special attention to the methods' implementation in modern computational tools.  The DOE, through its NEAMS (Nuclear Energy Advanced Modeling and Simulation) program, has especially sought the development of transient multiphysics capability in INL's MOOSE (Multiphysics Object-Oriented Simulation Environment).  The following two sections overview time-dependent neutronics and multiphysics solution methods.

\section{Time-Dependent Neutron Transport}

Transient modeling of nuclear reactors has been a chronically foreboding task due to its computationally expensive nature. The difficulty in transient reactor modeling is due to:
\begin{enumerate}
\item
the high-dimensionality of the phase-space for the governing equations that describe the flux of neutrons (6-D+time for multigroup neutron transport and 4-D+time for multigroup neutron diffusion), and 
\item
the fact that the time discretization has to be implicit, which  leads to stiff system of equations.  
\end{enumerate}
The transport equation has seven independent variables: space ($\vec{r}$), energy ($E$), direction ($\vec{\Omega}$), and time ($t$) \cite{duderstadt1976nuclear}.  Nuclear reactor simulations often utilize the neutron diffusion approximation, which carefully eliminates the dependence on direction.  For the purpose of this research, neutron diffusion will be discussed exclusively.

For computational purposes, the neutron diffusion equation is discretized in space, energy, and time.  There are several viable discretization schemes for each of these variables; FEM for space and multigroup in energy are used for the development of this research \cite{zeinkiewicz2005finite, duderstadt1976nuclear}. Since the purpose of this research involves temporal dependence, discretization in time is kept general. Most reactor computation frameworks discretize the time variable of flux directly using a multitude of schemes (Euler, Crank-Nicholson, Runge-Kutta, etc.).  However, the solutions of flux can be incredibly stiff due to the speed of neutrons, especially in a pulsing reactor \cite{TWIGL_benchmark}. Therefore, computationally expensive implicit schemes are necessary with many points of discretization (time steps). The following subsections describe two alternate techniques: point reactor kinetics and the improved quasi-static method (IQS). These methods are meant to reduce computational expense of evaluating transient flux while maintaining accuracy.

\subsection{Point Reactor Kinetics}

A common solution process for neutron kinetics is the evaluation of the point reactor kinetics equation (PRKE) \cite{Planchard_1991}.  The derivation of this equation involves factorization flux into a space-dependent shape and time-dependent amplitude.  This factorization immediately creates a large assumption that the spacial variance of the flux is time-independent.  The amplitude maintains the temporal stiffness as flux; however, the PRKE takes very little computational effort to evaluate because the multitudinous variables from the spacial discretization only needs to be evaluated on the initial time step.  The rest of the time steps only evaluated one variable: amplitude.

\subsection{Improved Quasi-Static Method}

IQS is a spatial kinetics method that involves factorizing the flux solution into space- and time-dependent components \cite{Ott_1966,Dulla2008,Devooght_1984,Monier_diss,Sissaoui_1995}.  These components are the flux's amplitude and its shape. Amplitude is only time-dependent, while the shape is both space- and time-dependent.  However, the impetus of the method is the assumption that the shape is only weakly dependent on time; therefore, the variable for the diffusion evaluation is less stiff without making the assumption that spacial variance is time-independent, as the PRKE does.  As a result, the shape may not require an update at the same frequency as the amplitude function, but only on larger macro-time steps.

Since shape and amplitude share a independent variable, they are temporally coupled. Ott in \cite{Ott_1966} was the first publication that investigated the coupling of shape and amplitude in a quasi-static nature, but kept steady-state shape equation to evaluate. Ott later in \cite{Ott_1969} incorporated the time derivative of shape in the equation, yielding better results.  The resulting system of equations is heavily nonlinear \cite{Dulla2008}. Nonlinear problems require an iterative process to evaluate, either fixed-point (Picard) or Newton iteration. Sissaoui et al. \cite{Sissaoui_1995}, Koclas et al. \cite{Koclas_1996}, Devooght et al. \cite{Devooght_1984}, and Monier \cite{Monier_diss} all use fixed-point iterative techniques for their quasi-static simulations, the main difference is their criteria for convergence.  Devooght et al. in \cite{Devooght_1984} also utilizes a Newton iteration technique.

\section{Time-Dependent Reactor Dynamics}

Transient reactor analysis also involves coupling neutronics with other physics like temperature feedback, thermal hydraulics, fuel structure and composition, etc., this coupling is known as multiphysics simulation. Multiphysics simulation is a daunting task because it requires communication of coupled variables that are being evaluated with a myriad of techniques. The coupling is also usually nonlinear requiring Picard or Newton iteration techniques. Programs, such as MOOSE, provide the capability of evaluating, communicating, and iterating different physics.

MOOSE is a multiphysics framework being developed at INL that presents the architecture for physics-based applications \cite{moose}.  Rattlesnake is a large application in MOOSE that involves deterministic radiation transport physics.  Currently, Rattlesnake is able to solve steady-state, transient, and k-eigenvalue neutron transport and diffusion problems using finite element methods (FEM) \cite{wang2013}. BISON is the application for evaluating fuel temperature and structural properties \cite{bison}.  RELAP-7 is the main safety analysis application, but focuses on variables involved with hydrodynamic analysis \cite{relap7}.  MAMMOTH is the over-arching application that couples all these physics by recomputing cross-sections for Rattlesnake and source terms for BISON and RELAP-7 \cite{mammoth}.

IQS is a manipulation of neutronics variables, so it is solely developed within the Rattlesnake application. However, proper treatment of multiphysics with IQS is vital for computational optimization.  The quasi-static nature of IQS can be extended to multiphysics simulation by determining the proper time scales for each physics.  Considering TREAT, shape is the slowest varying variable and amplitude is fastest, while fuel temperature is in between.  Varying time step sizes for each variable is optimal for minimizing computation expense, while maintaining accuracy.

\section{Objective}

The goal of this research is to continue the investigation and development of the improved quasi-static method.  This goal requires the development of IQS in a easily modifiable, prototype-like code, as well as a robust FEM framework. \\
\indent The prototyping code is used to investigate time discretization and nonlinear iteration schemes. The prototype will also guide the development of IQS in the large FEM framework of Rattlesnake/MOOSE, where more complex examples and benchmarks can be tested.  Development in Rattlesnake also spawns the opportunity to test IQS with multiphysics, namely temperature feedback, and time adaptation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tcr{
I think an outline for the introduction should be 
\begin{enumerate}
\item reactor transient testing. Background, etc... this subsection should conclude on the overarching need of transient testing: computational methods for multiphysics reactor transients. It should also introduce the next 2 subsections, which are both about lit review, one on time-dependent  neutronics, one on multiphysics solution techniques for reactor physics.
\item literature review on mostly neutronics time dependent: challenges in computational methods for time-dependent neutron balance equations. the size of the phase-space, the need for implicitness in the temporal treatment due to the stiffness, .... this should open up to a discussion on alternate methods, such as IQS, ... it is a lit review, so no need for equations, but rather a demonstration that you know the background and have a solid understand of that foundation to lay the ground from your OWN work. As part of this lit, some discussion of Mund's, Dulla's, Ott's, Monier,'s the Canadian's papers is welcome. 
\item now, reactor transients also means coupled to other physics. this is were you discuss some aspect related to multiphysics. you can start with the general discussion that this is typically a nonlinear problem, that it can be solved via picard or newton. this is maybe a segway into a discussion/introduction about MOOSE. Then you go back on IQS and add the multiphysics aspects. Ott did some work with multiphysics, I gave you the code manual of his own software where he discusses when he decided to do shape updates.
\end{enumerate}
Generally speaking, the rest is real good but you need to spend some time on these 5-8 pages of introduction that set the tone and the background. Everybody needs to understand that you know a lot; that your thoughts are well organized and clear; the rest will come easy (except some results ...).\\
side note: we need to put the small results about sdirk being poor for stiffness problem. maybe as an appendix. I want to keep that knowledge somewhere. 
}