### 2018-05-29 doc/Papers/IQS_convergence/highlights-iqs.tex
Adding highlights for submission

\documentclass{elsarticle}

\begin{document}

\section*{Highlights}

\begin{itemize}
\item Improved Quasi-Static Method (IQS) shows proper error convergence through fourth-order discretization, but performance diminishes compared to standard discretization.
\item IQS uniqueness criteria can only be preserved through a highly-accurate analytical treatment of delayed neutron precursors.
\item The effectiveness of IQS is maximized with time adaptation of shape updates.
\item Quasi-static treatment of other kinetic variables, such as temperature, significantly improves IQS performance for multi-physics problems.
\end{itemize}

\end{document}