### 2018-10-02 doc/Papers/IQS_convergence/reply_to_editor.tex
minor changes. splitting reply in 2 because I think you need to upload 2 letters

\documentclass[11pt]{letter}

%%%%%% Letter Size Setup %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
        \setlength{\textwidth}{6in}     %% For longer or shorter text width
        \setlength{\topmargin}{0in}    %% For more or less top margin
       \setlength{\textheight}{8.6in}    %% For longer or shorter textheight
        \setlength{\oddsidemargin}{0.25in} %% For odd side margin (twoside)
                                            %% or margin (oneside)
        \setlength{\paperwidth}{8.5in}
        \setlength{\paperheight}{11in}
        \setlength{\topmargin}{0pt}
        \setlength{\headheight}{0pt}
        \setlength{\headsep}{0pt}
        \setlength{\footskip}{0.4in}
        \setlength{\marginparwidth}{0in}
        \setlength{\marginparsep}{0in}
        \setlength{\hoffset}{0in}
        \setlength{\voffset}{0in}

\address{Zachary M. Prince\\ 
Department of Nuclear Engineering \\
Texas A\&M University\\
College Station, TX 77843-3133, USA\\
phone: (360) 362 6622\\
e-mail: zachmprince@tamu.edu \vspace{0.5cm}}

%%%%%% The Signature  and Date %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\signature{\vspace{-1.25cm}Zachary M. Prince, Jean C. Ragusa}

\usepackage{xspace}

\newcommand{\iqspc}{IQS-PC\xspace}
\newcommand{\section}[1]{\textbf{#1}}

%\newcommand{\fix}{\textsquare}
%\newcommand{\working}{$\boxdot$}
%\newcommand{\nofix}{?}
%\newcommand{\done}{\checkmark}
\newcommand{\fix}{$\bullet$}
\newcommand{\working}{$\bullet$}
\newcommand{\nofix}{$\bullet$}
\newcommand{\done}{$\bullet$}

%\newcommand{\easy}[1]{\textcolor{ForestGreen}{#1}}
%\newcommand{\medm}[1]{\textcolor{BurntOrange}{#1}}
%\newcommand{\hard}[1]{\textcolor{Red}{#1}}
\newcommand{\easy}[1]{{\textit{#1}}}
\newcommand{\medm}[1]{{\textit{#1}}}
\newcommand{\hard}[1]{{\textit{#1}}}

%\usepackage{color}
%\newcommand{\tcr}[1]{\textcolor{red}{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%

\begin{letter}{Professor Sara A. Pozzi\\
    Editor,\\
    Elsevier Annals of Nuclear Energy\\
    \bigskip
    Re: Revision of manuscript ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method'', ANUCENE-D-18-00462}
\date{\today}
%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%         \vspace{3.0cm}

\opening{Dear Professor Pozzi,}
         \vspace{0.25cm}
%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thank you for the opportunity to revise our manuscript, ``Multiphysics Reactor-core Simulations Using the Improved Quasi-Static Method''. We appreciate the prompt review and constructive suggestions by the reviewer. We believe our manuscript is substantially improved after making the suggested edits.

Overall, the comments from the reviewer were positive and pivotal in the improved quality of the manuscript. We are following a reply to the reviewer where the review comments are in italics with our responses below, including how and where the manuscript was modified.

We hope you will find the revised manuscript suitable for publication. Thank you again for considering this manuscript for publication in the {\it Annals of Nuclear Energy}.

\vspace{0.25cm}

%%%%%% More vertical space can be added here %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%% The Closing %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\closing{Sincerely, }

\end{letter}

\end{document}